### 2024-10-11 paper/paper.qmd
Logistic Regression added

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(boot)
library(broom.mixed)
library(collapse)
library(dataverse)
library(gutenbergr)
library(janitor)
library(knitr)
library(marginaleffects)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)
df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))

```

```{r}
#| warning: false
#| message: false
filtered_df <- df |> filter(year==2021) |> filter(age!=996) |> filter(marstat!=99) |> filter(!educ %in% c(996,997,998,999))
labeled_df <- filtered_df |> mutate(
  label_outcome = if_else(outcome==1, "Yes", "No"),
  label_outcome = as_factor(label_outcome),
  label_sex = if_else(sex==1,"Male", "Female"),
  label_education = case_when(
      educ==100~"Grade 12 or less, no high school diploma or equivalent",
      educ==101~"Grade 8 or less (no further detail)",
      educ==102~"Never attended/kindergarten only",
      educ==103~"Grades 1-11 (no further detail)",
      educ==104~"Grade 1",
      educ==105~"Grade 2",
      educ==106~"Grade 3",
      educ==107~"Grade 4",
      educ==108~"Grade 5",
      educ==109~"Grade 6",
      educ==110~"Grade 7",
      educ==111~"Grade 8",
      educ==112~"Grade 9-12, no diploma (no further detail)",
      educ==113~"Grade 9",
      educ==114~"Grade 10",
      educ==115~"Grade 11",
      educ==116~"12th grade, no diploma",
      educ==200~"High school diploma or GED",
      educ==201~"High school graduate",
      educ==202~"GED or equivalent",
      educ==300~"Some college, no 4yr degree",
      educ==301~"Some college, no degree",
      educ==302~"AA degree: technical/vocational/occupational",
      educ==303~"AA degree: academic program",
      educ==400~"Bachelor's degree (BA,AB,BS,BBA)",
      educ==500~"Master's, Professional, or Doctoral Degree",
      educ==501~"Master's degree (MA,MS,Med,MBA)",
      educ==502~"Professional (MD,DDS,DVM,JD)",
      educ==503~"Doctoral degree (PhD, EdD)",
      educ==504~"Other degree",
      educ==505~"Professional School or Doctoral degree, topcoded (MD, DDS, DVM, JD, PhD, EdD)"
  ),
  label_education = factor(
    label_education,
    levels=c(
      "Grade 12 or less, no high school diploma or equivalent",
      "Grade 8 or less (no further detail)",
      "Never attended/kindergarten only",
      "Grades 1-11 (no further detail)",
      "Grade 1",
      "Grade 2",
      "Grade 3",
      "Grade 4",
      "Grade 5",
      "Grade 6",
      "Grade 7",
      "Grade 8",
      "Grade 9-12, no diploma (no further detail)",
      "Grade 9",
      "Grade 10",
      "Grade 11",
      "12th grade, no diploma",
      "High school diploma or GED",
      "High school graduate",
      "GED or equivalent",
      "Some college, no 4yr degree",
      "Some college, no degree",
      "AA degree: technical/vocational/occupational",
      "AA degree: academic program",
      "Bachelor's degree (BA,AB,BS,BBA)",
      "Master's, Professional, or Doctoral Degree",
      "Master's degree (MA,MS,Med,MBA)",
      "Professional (MD,DDS,DVM,JD)",
      "Doctoral degree (PhD, EdD)",
      "Other degree",
      "Professional School or Doctoral degree, topcoded (MD, DDS, DVM, JD, PhD, EdD)"
    )
  )
)

labeled_df |>
  ggplot(aes(x = label_education, fill = label_outcome)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(label_sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Mental Depression"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "bottom")
```

```{r}
#| warning: false
#| message: false
labeled_df_reduced <- labeled_df |> slice_sample(n = 1000)
mental_illness <-
  stan_glm(
    label_outcome ~ label_sex + label_education,
    data = labeled_df_reduced,
    family = binomial(link = "logit"),
    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
    prior_intercept = 
      normal(location = 0, scale = 2.5, autoscale = TRUE),
    seed = 853
  )
saveRDS(
  mental_illness,
  file = here::here("models/mental_illness.rds")
)
```
```{r}
#| warning: false
#| message: false
mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))

modelsummary(
  list(
    "Mental Illness Exist" = mental_illness
  ),
  statistic = "mad"
  )
```
# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-11 scripts/03-clean_data.R

@@ -1,44 +1,24 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data - removed rows where K6SUM==98,96 
+# Author: Marcus Barnes | Christina Nguyen | Nibras Ar Rakib
+# Date: 11 Oct 2024
+# Contact: nibras.rakib@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download raw data from https://meps.ipums.org/meps/
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/01-raw_data/meps_00003.csv")
 
 cleaned_data <-
   raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+  janitor::clean_names()
 
+cleaned_data <- cleaned_data |> select(year,pid,age,sex,marstat,educ,inctot,k6sum)
+
[…]

### 2024-10-11 paper/paper.qmd
Added code for model generation

@@
 #| message: false
-filtered_df <- df |> filter(year==2021) |> filter(age!=996) |> filter(marstat!=99) |> filter(!educ %in% c(996,997,998,999))
-labeled_df <- filtered_df |> mutate(
-  label_outcome = if_else(outcome==1, "Yes", "No"),
-  label_outcome = as_factor(label_outcome),
-  label_sex = if_else(sex==1,"Male", "Female"),
-  label_education = case_when(
-      educ==100~"Grade 12 or less, no high school diploma or equivalent",
-      educ==101~"Grade 8 or less (no further detail)",
-      educ==102~"Never attended/kindergarten only",
-      educ==103~"Grades 1-11 (no further detail)",
-      educ==104~"Grade 1",
-      educ==105~"Grade 2",
-      educ==106~"Grade 3",
-      educ==107~"Grade 4",
-      educ==108~"Grade 5",
-      educ==109~"Grade 6",
-      educ==110~"Grade 7",
-      educ==111~"Grade 8",
-      educ==112~"Grade 9-12, no diploma (no further detail)",
-      educ==113~"Grade 9",
-      educ==114~"Grade 10",
-      educ==115~"Grade 11",
-      educ==116~"12th grade, no diploma",
-      educ==200~"High school diploma or GED",
-      educ==201~"High school graduate",
-      educ==202~"GED or equivalent",
-      educ==300~"Some college, no 4yr degree",
-      educ==301~"Some college, no degree",
-      educ==302~"AA degree: technical/vocational/occupational",
-      educ==303~"AA degree: academic program",
-      educ==400~"Bachelor's degree (BA,AB,BS,BBA)",
-      educ==500~"Master's, Professional, or Doctoral Degree",
-      educ==501~"Master's degree (MA,MS,Med,MBA)",
-      educ==502~"Professional (MD,DDS,DVM,JD)",
-      educ==503~"Doctoral degree (PhD, EdD)",
-      educ==504~"Other degree",
-      educ==505~"Professional School or Doctoral degree, topcoded (MD, DDS, DVM, JD, PhD, EdD)"
-  ),
-  label_education = factor(
-    label_education,
-    levels=c(
-      "Grade 12 or less, no high school diploma or equivalent",
-      "Grade 8 or less (no further detail)",
-      "Never attended/kindergarten only",
-      "Grades 1-11 (no further detail)",
-      "Grade 1",
-      "Grade 2",
-      "Grade 3",
-      "Grade 4",
-      "Grade 5",
-      "Grade 6",
-      "Grade 7",
-      "Grade 8",
-      "Grade 9-12, no diploma (no further detail)",
-      "Grade 9",
-      "Grade 10",
-      "Grade 11",
-      "12th grade, no diploma",
-      "High school diploma or GED",
-      "High school graduate",
-      "GED or equivalent",
-      "Some college, no 4yr degree",
-      "Some college, no degree",
-      "AA degree: technical/vocational/occupational",
-      "AA degree: academic program",
-      "Bachelor's degree (BA,AB,BS,BBA)",
-      "Master's, Professional, or Doctoral Degree",
-      "Master's degree (MA,MS,Med,MBA)",
-      "Professional (MD,DDS,DVM,JD)",
-      "Doctoral degree (PhD, EdD)",
-      "Other degree",
-      "Professional School or Doctoral degree, topcoded (MD, DDS, DVM, JD, PhD, EdD)"
-    )
-  )
-)
-
 labeled_df |>
@@
 ```
-
-```{r}
-#| warning: false
-#| message: false
-labeled_df_reduced <- labeled_df |> slice_sample(n = 1000)
-mental_illness <-
-  stan_glm(
-    label_outcome ~ label_sex + label_education,
-    data = labeled_df_reduced,
-    family = binomial(link = "logit"),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = 
-      normal(location = 0, scale = 2.5, autoscale = TRUE),
-    seed = 853
-  )
-saveRDS(
-  mental_illness,
-  file = here::here("models/mental_illness.rds")
-)
-```
 ```{r}
@@
 ```
+
 # Introduction
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-10-11 scripts/06-model_data.R

@@ -1,37 +1,122 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Linear Regression
+# Author: Nibras Ar Rakib
+# Date: 11 October 2024
+# Contact: nibras.rakib@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 
 
 #### Workspace setup ####
-library(tidyverse)
+library(boot)
+library(broom.mixed)
+library(collapse)
+library(dataverse)
+library(gutenbergr)
+library(janitor)
+library(knitr)
+library(marginaleffects)
+library(modelsummary)
 library(rstanarm)
+library(tidybayes)
+library(tidyverse)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+df <- read.csv("data/02-analysis_data/analysis_data.csv")
+
+filtered_df <- df |> filter(year==2021) |> filter(age!=996) |> filter(marstat!=99) |> filter(!educ %in% c(996,997,998,999))
+
+labeled_df <- filtered_df |> mutate(
+  label_outcome = if_else(outcome==1, "Yes", "No"),
+  label_outcome = as_factor(label_outcome),
+  label_sex = if_else(sex==1,"Male", "Female"),
+  label_education = case_when(
+    educ==100~"Grade 12 or less, no high school diploma or equivalent",
+    educ==101~"Grade 8 or less (no further detail)",
+    educ==102~"Never attended/kindergarten only",
+    educ==103~"Grades 1-11 (no further detail)",
+    educ==104~"Grade 1",
+    educ==105~"Grade 2",
+    educ==106~"Grade 3",
+    educ==107~"Grade 4",
[…]

### 2024-10-17 paper/paper.qmd
Year 2018|2022 - analysis complete.

@@
 library(tidyverse)
-df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+labeled_df <- read.csv(here::here("data/02-analysis_data/model_data_sample_2k.csv"))
+```
+# Introduction
 
-```
+Overview paragraph
+
+Estimand paragraph
+
+Results paragraph
+
+Why it matters paragraph
+
+Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+
+# Data {#sec-data}
 
 ```{r}
+#| label: fig-education
+#| fig-cap: Education
 #| warning: false
 #| message: false
+#| echo: false
 labeled_df |>
@@
 ```
-```{r}
-#| warning: false
-#| message: false
-mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
-
-modelsummary(
-  list(
-    "Mental Illness Exist" = mental_illness
-  ),
-  statistic = "mad"
-  )
-```
-
-# Introduction
-
-Overview paragraph
-
-Estimand paragraph
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
-
-# Data {#sec-data}
-
 ## Overview
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# ggplot(penguins, aes(x = island, fill = species)) +
+#   geom_bar(alpha = 0.8) +
+#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
+#                     guide = "none") +
+#   theme_minimal() +
+#   facet_wrap(~species, ncol = 1) +
+#   coord_flip()
 ```
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# 
+# analysis_data |> 
+#   ggplot(aes(x = width, y = length)) +
+#   geom_point(alpha = 0.8) +
+#   theme_minimal() +
+#   labs(x = "Wing width (mm)",
+#        y = "Wing length (mm)")
 ```
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-lrmodelresults.
 
 ```{r}
+#| warning: false
+#| message: false
 #| echo: false
 #| eval: true
-#| warning: false
-#| message: false
+#| label: tbl-lrmodelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+modelsummary(
+  list(
+    "Mental Illness Exist" = mental_illness
+  ),
+  statistic = "mad"
+  )
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# modelsummary::modelsummary(
+#   list(
+#     "First model" = first_model
+#   ),
+#   statistic = "mad",
+#   fmt = 2
+# )
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+# pp_check(first_model) +
+#   theme_classic() +
+#   theme(legend.position = "bottom")
+# 
+# posterior_vs_prior(first_model) +
+#   theme_minimal() +
+#   scale_color_brewer(palette = "Set1") +
+#   theme(legend.position = "bottom") +
+#   coord_flip()
 ```
@@
 #| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
+# 
+# plot(first_model, "trace")
+# 
+# plot(first_model, "rhat")
 ```

### 2024-10-17 scripts/03-clean_data.R

@@ -16,7 +16,7 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names()
 
-cleaned_data <- cleaned_data |> select(year,pid,age,sex,marstat,educ,inctot,k6sum)
+cleaned_data <- cleaned_data |> select(year,pid,age,sex,marstat,racea,educ,inctot,k6sum)
 
 cleaned_data <- cleaned_data |> filter(!k6sum %in% c(96,98)) |> filter(inctot>0)
 cleaned_data$outcome <- ifelse(cleaned_data$k6sum<13,0,1)

### 2024-10-17 scripts/05-exploratory_data_analysis.R

@@ -1,37 +1,498 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Regression
+# Author: Nibras Ar Rakib
+# Date: 15 October 2024
+# Contact: nibras.rakib@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
-library(tidyverse)
+library(boot)
+library(broom.mixed)
+library(collapse)
+library(dataverse)
+library(gutenbergr)
+library(janitor)
+library(knitr)
+library(marginaleffects)
+library(modelsummary)
 library(rstanarm)
+library(tidybayes)
+library(tidyverse)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+
+#### Read data ####
+
+df <- read.csv("data/02-analysis_data/analysis_data.csv")
+
+### Model data ###
+filtered_df_2018 <- df |> filter(year==2018) |> filter(age!=996) |>
+  filter(age>18) |> filter(age<80) |>
+  filter(sex %in% c(1,2)) |> filter(marstat %in% c(10,20,30,40,50,99)) |>
+  filter(racea %in% c(100,200,300,310,320,330,340,350,400,410,411,412,413,414,415,416,420,421,422,423,430,431,432,433,434,500,510,520,530,540,550,560,570,580,600,610,611,612,613,614,615,616,617)) |> 
[…]

### 2024-10-17 scripts/06-model_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Linear Regression
+# Purpose: Regression
 # Author: Nibras Ar Rakib
 # Date: 11 October 2024
 # Contact: nibras.rakib@mail.utoronto.ca
@@ -20,103 +20,4 @@ library(rstanarm)
 library(tidybayes)
 library(tidyverse)
 
-#### Read data ####
-
-df <- read.csv("data/02-analysis_data/analysis_data.csv")
-
-filtered_df <- df |> filter(year==2021) |> filter(age!=996) |> filter(marstat!=99) |> filter(!educ %in% c(996,997,998,999))
-
-labeled_df <- filtered_df |> mutate(
-  label_outcome = if_else(outcome==1, "Yes", "No"),
-  label_outcome = as_factor(label_outcome),
-  label_sex = if_else(sex==1,"Male", "Female"),
-  label_education = case_when(
-    educ==100~"Grade 12 or less, no high school diploma or equivalent",
-    educ==101~"Grade 8 or less (no further detail)",
-    educ==102~"Never attended/kindergarten only",
-    educ==103~"Grades 1-11 (no further detail)",
-    educ==104~"Grade 1",
-    educ==105~"Grade 2",
-    educ==106~"Grade 3",
-    educ==107~"Grade 4",
-    educ==108~"Grade 5",
-    educ==109~"Grade 6",
-    educ==110~"Grade 7",
-    educ==111~"Grade 8",
-    educ==112~"Grade 9-12, no diploma (no further detail)",
-    educ==113~"Grade 9",
-    educ==114~"Grade 10",
-    educ==115~"Grade 11",
-    educ==116~"12th grade, no diploma",
-    educ==200~"High school diploma or GED",
-    educ==201~"High school graduate",
-    educ==202~"GED or equivalent",
-    educ==300~"Some college, no 4yr degree",
-    educ==301~"Some college, no degree",
-    educ==302~"AA degree: technical/vocational/occupational",
-    educ==303~"AA degree: academic program",
-    educ==400~"Bachelor's degree (BA,AB,BS,BBA)",
-    educ==500~"Master's, Professional, or Doctoral Degree",
-    educ==501~"Master's degree (MA,MS,Med,MBA)",
-    educ==502~"Professional (MD,DDS,DVM,JD)",
-    educ==503~"Doctoral degree (PhD, EdD)",
-    educ==504~"Other degree",
[…]

### 2024-10-21 paper/paper.qmd
Adds Chritina's initial text suggestions and removes some unnecessary filler from the starter_folder.

--- 2024-10-11 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Gender, Education, and Mental Depression"
+subtitle: "Analyzing Census Data Through Statistical Methods"
 author: 
-  - First author
-  - Another author
+  - Nibras Ar Rakib
+  - Christina Nguyen
+  - Marcus Emmanuel Barnes
 thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
@@
 library(tidyverse)
-df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
+# df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+labeled_df <- read.csv(here::here("data/02-analysis_data/model_data_sample_2k.csv"))
 ```
-
-```{r}
-#| warning: false
-#| message: false
-labeled_df |>
-  ggplot(aes(x = label_education, fill = label_outcome)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(label_sex)) +
-  theme_minimal() +
-  labs(
-    x = "Highest education",
-    y = "Number of respondents",
-    fill = "Mental Depression"
-  ) +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set2") +
-  theme(legend.position = "bottom")
-```
-```{r}
-#| warning: false
-#| message: false
-mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
-
-modelsummary(
-  list(
-    "Mental Illness Exist" = mental_illness
-  ),
-  statistic = "mad"
-  )
-```
-
 # Introduction
+Key questions include: Does higher education levels correlate with lower rates of depression? Are women disproportionately affected by depression compared to men at the same educational level?
 
@@
 
+The data for this study is drawn from national census data, which includes variables for gender (male/female), education level (from grade 1 to doctoral degree), and mental depression (no/yes).
+The sample consists of individuals age (FILL THIS IN AND WHY WE CHOSE THIS).
+
+```{r}
+#| label: fig-education
+#| fig-cap: Education
+#| warning: false
+#| message: false
+#| echo: false
+labeled_df |>
+  ggplot(aes(x = education, fill = outcome)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(sex)) +
+  theme_minimal() +
+  labs(
+    x = "Highest education",
+    y = "Number of respondents",
+    fill = "Mental Depression"
+  ) +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set2") +
+  theme(legend.position = "bottom")
+```
+The data for this study is drawn from national census data, which includes variables for gender (male/female), education level (from grade 1 to doctoral degree), and mental depression (no/yes).
+The sample consists of individuals age (FILL THIS IN AND WHY WE CHOSE THIS).
+
 ## Overview
@@
 
-Overview text
+### Gender and Mental Health
+Research consistently shows that women report higher rates of depression than men (CITE).
+Biological, social, and psychological factors contribute to this disparity.
+Women may experience unique stressors, such as role strain from balancing work and family obligations, as well as being subject to gender-based discrimination (CITE).
+These pressures can exacerbate mental health issues, leading to higher rates of depression among women, particularly those in lower socioeconomic or educational strata.
+
+### Educational Attainment and Mental Health
+Education is often seen as a protective factor against mental health issues.
+Studies suggest that higher levels of education correlate with improved mental health outcomes (CITE).
+Educated individuals tend to have better access to mental health resources, higher socioeconomic status, and greater job security, all of which contribute to lower rates of depression (CITE).
+(DO WE WANT DATA ON HEALTH CARE COVERAGE/INSURANCE HERE?) Conversely, those with lower education levels may face more financial strain, reduced employment opportunities, and limited access to healthcare, increasing their vulnerability to depression.
+
+### Intersection of Gender, Education, and Depression
+Despite the protective nature of education, the gender gap in mental health persists even at higher levels of educational attainment.
+Some studies suggest that women with high education levels still experience depression due to external stressors, including workplace discrimination and societal expectations (CITE).
+Additionally, the mental health benefits of education may differ between genders, with men benefiting more from the social status and economic opportunities that education provides, while women may continue to face challenges related to gender roles (cite).
+
+## Methodology
+
+### Statistical Techniques
+This study employs the following statistical techniques:
+
+* Descriptive statistics : These will summarize the distribution of education levels and depression rates across genders.
+* Logistic regression: We will use logistic regression to examine how education level and gender predict the likelihood of experiencing depression, controlling for other demographic factors like age and income.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+Some paragraphs about how we go from a phenomena in the world to an entry in the dataset (that is talk about the data was collected).
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
 Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
 
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+We expect a positive relationship between education level and
 
@@
 
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
 # Discussion
 
-## First discussion point {#sec-first-point}
+## Gender Differences in Depression {#discussion-first-point}
+The finding that women report higher rates of depression across all educational levels aligns with existing literature.
+This disparity can be attributed to a variety of factors, including biological differences in stress response and the disproportionate burden women face in managing work-life balance (CITE).
+Additionally, women may be more likely to seek help and report depressive symptoms, contributing to higher rates in survey data (CITE).
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Education as a Protective Factor {#discussion-second-point}
+The results support the hypothesis that education serves as a protective factor against depression.
+Those with higher education levels enjoy greater access to economic and social resources, reducing the risk of depression (CITE).
+However, the persistent gender gap in mental health, even among those with higher education, suggests that education alone is insufficient to fully mitigate the effects of gender-based stressors and societal pressures on women (CITE AND CHECK MY FACTS HERE).
 
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
+## Intersectionality and Mental Health
+The intersection of gender and education highlights the complexity of mental health outcomes.
+Women with lower education levels are doubly disadvantaged, experiencing both the socioeconomic limitations of low education and the additional burdens associated with gender.
+Future research should explore how other intersecting factors, such as race and socioeconomic status, further influence the relationship between education and mental health.
 
@@
 Weaknesses and next steps should also be included.
+
+# Conclusion
+This paper has demonstrated that both education and gender play crucial roles in shaping mental health outcomes, particularly depression.[ETC FILL THIS IN]
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-10-21 paper/paper.qmd
added marcus's writeup

--- 2024-10-17 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Gender and Mental Depression: Analyzing Census Data Through Statistical Methods"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Christina Nguyen
+  - Marcus Barnes
+  - Nibras Ar Rakib
+thanks: "Code and data are available at: [https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty](https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty)."
 date: today
@@
 library(tidyverse)
-# df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
-labeled_df <- read.csv(here::here("data/02-analysis_data/model_data_sample_2k.csv"))
+library(here)
+filtered_df_2018 <- read.csv(here::here("data/02-analysis_data/model_data_2018.csv"))
+filtered_df_2021 <- read.csv(here::here("data/02-analysis_data/model_data_2021.csv"))
 ```
+
 # Introduction
@@
 #| echo: false
-labeled_df |>
-  ggplot(aes(x = label_education, fill = label_outcome)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(label_sex)) +
-  theme_minimal() +
-  labs(
-    x = "Highest education",
-    y = "Number of respondents",
-    fill = "Mental Depression"
-  ) +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set2") +
-  theme(legend.position = "bottom")
+# labeled_df |>
+#   ggplot(aes(x = label_education, fill = label_outcome)) +
+#   stat_count(position = "dodge") +
+#   facet_wrap(facets = vars(label_sex)) +
+#   theme_minimal() +
+#   labs(
+#     x = "Highest education",
+#     y = "Number of respondents",
+#     fill = "Mental Depression"
+#   ) +
+#   coord_flip() +
+#   scale_fill_brewer(palette = "Set2") +
+#   theme(legend.position = "bottom")
 ```
+
 ## Overview
+We used the Medical Expenditure Panel Survey dataset from the Integrated Public Use Microdata Series, which contains microdata from different time points related to U.S.
+healthcare expenditures [@blewett_ipums_2023].
+We performed all analysis using a popular statistical language called R [@citeR].
+Our dataset contains age, sex, marital status, race, highest level of education attainment, total personal income, and K6 score [@kessler_kessler_2012].All the variable definitions can be found in the Appendix [-@sec-appendix-data-details].
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+Our goal was to identify people, based on gender and education, who are suffering from non-specific psychological distress based on the Kessler scale [@kessler_kessler_2012] before and after the COVID-19 pandemic.
 
-Overview text
-
+Our goal was to identify respondents, based on gender and education, who were suffering from non-specific psychological distress based on the Kessler scale before and after the COVID-19 pandemic.
+To support our analyses, we split into two datasets for 2018 and 2021.
+We also excluded the respondents whose ages are below 19 and over 80.
+After enforcing our inclusion criteria, we had 15716 respondents for 2018 and 12685 for 2021.
+To give a preview of our data, @tbl-2018 and @tbl-2021 show the first five rows for each row for 2018 and 2021, respectively.
+```{r}
+#| label: fig-education-number-rows
+#| fig-cap: Education
+#| warning: false
+#| message: false
+#| echo: false
+#| results: hide
+print(nrow(filtered_df_2018))
+print(nrow(filtered_df_2021))
+```
+```{r}
+#| label: tbl-2018
+#| tbl-cap: "Year 2018"
+#| warning: false
+#| message: false
+#| echo: false
+kable(filtered_df_2018[1:5,c("year","age","sex","marstat","racea","educ","inctot","k6sum")])
+```
+```{r}
+#| label: tbl-2021
+#| tbl-cap: "Year 2021"
+#| warning: false
+#| message: false
+#| echo: false
+kable(filtered_df_2021[1:5,c("year","age","sex","marstat","racea","educ","inctot","k6sum")])
+```
 ## Measurement
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+### Gender and Mental Health
+
+Research consistently shows that women report higher rates of depression than men (CITE).
+Biological, social, and psychological factors contribute to this disparity.
+Women may experience unique stressors, such as role strain from balancing work and family obligations, as well as being subject to gender-based discrimination (CITE).
+These pressures can exacerbate mental health issues, leading to higher rates of depression among women, particularly those in lower socioeconomic or educational strata.
+
+### Educational Attainment and Mental Health
+
+Education is often seen as a protective factor against mental health issues.
+Studies suggest that higher levels of education correlate with improved mental health outcomes (CITE).
+Educated individuals tend to have better access to mental health resources, higher socioeconomic status, and greater job security, all of which contribute to lower rates of depression (CITE).
+Conversely, those with lower education levels may face more financial strain, reduced employment opportunities, and limited access to healthcare, increasing their vulnerability to depression.
+
+### Intersection of Gender, Education, and Depression
+
+Despite the protective nature of education, the gender gap in mental health persists even at higher levels of educational attainment.
+Some studies suggest that women with high education levels still experience depression due to external stressors, including workplace discrimination and societal expectations (CITE).
+Additionally, the mental health benefits of education may differ between genders, with men benefiting more from the social status and economic opportunities that education provides, while women may continue to face challenges related to gender roles (cite).
 
@@
 #| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
-
-modelsummary(
-  list(
-    "Mental Illness Exist" = mental_illness
-  ),
-  statistic = "mad"
-  )
+# mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
+# 
+# modelsummary(
+#   list(
+#     "Mental Illness Exist" = mental_illness
+#   ),
+#   statistic = "mad"
+#   )
 ```
@@
 
-# Additional data details
+# Additional data details {#sec-appendix-data-details}
 
+## Variable Definitions
+Variable: "AGE"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">AGE</td></tr><tr><td class="label">Label:</td><td class="value">Age</td></tr><tr><td class="label">Variable Text:</td><td class="value">AGE reports the individual's exact age, calculated from date of birth, as of the last day (12/31) of the survey year.
+<br>&amp;nbsp;Date
+ of birth and age were asked for each reporting unit member, and then 
+exact age was calculated from date of birth. Where the calculated age 
+and the age provided did not match, inconsistencies were reviewed and 
+resolved. When date of birth was not provided, but age was provided, the
+ month and year of birth were assigned randomly from among the possible 
+valid options.
+For any cases still not accounted, age was imputed using:<br>(1) the mean age difference between MEPS participants with certain family relationships (where available) or  <br>(2) the mean age value for MEPS participants.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">99</td></tr><tr><td class="label">End Position:</td><td class="value">101</td></tr><tr><td class="label">Width:</td><td class="value">3</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="label">Coder Instructions:</td><td class="value">CodesAGE is a 3-digit-numeric variable.<br><br>085:	Top code for 85 years or older (2001-forward)<br>090:	Top code for 90 years or older (1996-2000)<br>996:	Missing</td></tr></tbody></table>
+
+```
+Variable: "Sex"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">SEX</td></tr><tr><td class="label">Label:</td><td class="value">Sex</td></tr><tr><td class="label">Variable Text:</td><td class="value">SEX indicates whether the person was male or female.
+<br><br>Collection
+ of information on SEX in MEPS&amp;nbsp;Data on the sex of each 
+reporting unit (RU) member was determined during the NHIS interview, and
+ was then verified, and, if necessary, corrected during each MEPS 
+interview.<br><br>If the respondent was a new RU member or their sex was
+ not ascertained in the NHIS interview, MEPS initially used the first 
+name of the respondent to assign their sex. If the first name gave no 
+clear indication of sex, the reported family relationships were used to 
+assign sex. If the sex of the respondent was still unclear, sex was 
+randomly assigned. <br><br>The NHIS method of ascertaining the sex of 
+the respondent, which primarily informs the MEPS version of SEX, is 
+similar to the MEPS method. First, sex of the respondent was inferred 
+from the individual's first name or family relationships. If the sex of 
+the respondent was unclear, the interviewer was instructed to explicitly
+ ask the person's sex. Beginning in 1998, interviewers were told to 
+"enter your best guess" when the respondent either did not know or 
+refused to answer the direct question that was asked about the person's 
+sex.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">102</td></tr><tr><td class="label">End Position:</td><td class="value">102</td></tr><tr><td class="label">Width:</td><td class="value">1</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
+<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">1</td><td class="cat_lbl">Male</td></tr><tr><td class="cat_val">2</td><td class="cat_lbl">Female</td></tr><tr><td class="cat_val">7</td><td class="cat_lbl">Unknown-refused</td></tr><tr><td class="cat_val">8</td><td class="cat_lbl">Unknown-not ascertained</td></tr><tr><td class="cat_val">9</td><td class="cat_lbl">Unknown-don't know</td></tr></tbody></table>
+```
+Variable: "MARSTAT"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">MARSTAT</td></tr><tr><td class="label">Label:</td><td class="value">Legal marital status</td></tr><tr><td class="label">Variable Text:</td><td class="value">For persons age 16 and older, MARSTAT reports the person's legal marital status.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">103</td></tr><tr><td class="label">End Position:</td><td class="value">104</td></tr><tr><td class="label">Width:</td><td class="value">2</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
+<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">00</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">10</td><td class="cat_lbl">Married</td></tr><tr><td class="cat_val">20</td><td class="cat_lbl">Widowed</td></tr><tr><td class="cat_val">30</td><td class="cat_lbl">Divorced</td></tr><tr><td class="cat_val">40</td><td class="cat_lbl">Separated</td></tr><tr><td class="cat_val">50</td><td class="cat_lbl">Never married</td></tr><tr><td class="cat_val">99</td><td class="cat_lbl">Unknown marital status</td></tr></tbody></table>
+```
+Variable: "RACEA"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">RACEA</td></tr><tr><td class="label">Label:</td><td class="value">Main Racial Background (Pre-1997 Revised OMB Standards), self-reported or interviewer reported</td></tr><tr><td class="label">Variable Text:</td><td class="value">RACEA
+ reports the race of the respondent. If not ascertained, the race and/or
+ ethnicity were assigned based on relationship to other members of the 
+DU using a priority ordering that gave precedence to blood relatives in 
+the immediate family.</td></tr><tr><td class="label">Concept:</td><td class="value">Ethnicity/Nativity Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">105</td></tr><tr><td class="label">End Position:</td><td class="value">107</td></tr><tr><td class="label">Width:</td><td class="value">3</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
+<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">100</td><td class="cat_lbl">White</td></tr><tr><td class="cat_val">200</td><td class="cat_lbl">Black/African-American</td></tr><tr><td class="cat_val">300</td><td class="cat_lbl">Aleut, Alaskan Native, or American Indian</td></tr><tr><td class="cat_val">310</td><td class="cat_lbl">Alaskan Native or American Indian</td></tr><tr><td class="cat_val">320</td><td class="cat_lbl">Alaskan Native/Eskimo</td></tr><tr><td class="cat_val">330</td><td class="cat_lbl">Aleut</td></tr><tr><td class="cat_val">340</td><td class="cat_lbl">American Indian</td></tr><tr><td class="cat_val">350</td><td class="cat_lbl">American Indian or Alaskan Native and any other group</td></tr><tr><td class="cat_val">400</td><td class="cat_lbl">Asian or Pacific Islander</td></tr><tr><td class="cat_val">410</td><td class="cat_lbl">Asian</td></tr><tr><td class="cat_val">411</td><td class="cat_lbl">Chinese</td></tr><tr><td class="cat_val">412</td><td class="cat_lbl">Filipino</td></tr><tr><td class="cat_val">413</td><td class="cat_lbl">Korean</td></tr><tr><td class="cat_val">414</td><td class="cat_lbl">Vietnamese</td></tr><tr><td class="cat_val">415</td><td class="cat_lbl">Japanese</td></tr><tr><td class="cat_val">416</td><td class="cat_lbl">Asian Indian</td></tr><tr><td class="cat_val">420</td><td class="cat_lbl">Pacific Islander</td></tr><tr><td class="cat_val">421</td><td class="cat_lbl">Hawaiian</td></tr><tr><td class="cat_val">422</td><td class="cat_lbl">Samoan</td></tr><tr><td class="cat_val">423</td><td class="cat_lbl">Guamanian</td></tr><tr><td class="cat_val">430</td><td class="cat_lbl">Other Asian or Pacific Islander</td></tr><tr><td class="cat_val">431</td><td class="cat_lbl">Other Asian or Pacific Islander (1992-1995)</td></tr><tr><td class="cat_val">432</td><td class="cat_lbl">Other Asian or Pacific Islander (1996)</td></tr><tr><td class="cat_val">433</td><td class="cat_lbl">Other Asian or Pacific Islander (1997-1998)</td></tr><tr><td class="cat_val">434</td><td class="cat_lbl">Other Asian (1999 forward)</td></tr><tr><td class="cat_val">500</td><td class="cat_lbl">Other Race</td></tr><tr><td class="cat_val">510</td><td class="cat_lbl">Other Race (1963-1977)</td></tr><tr><td class="cat_val">520</td><td class="cat_lbl">Other Race (1978)</td></tr><tr><td class="cat_val">530</td><td class="cat_lbl">Other Race (1979-1991)</td></tr><tr><td class="cat_val">540</td><td class="cat_lbl">Other Race (1992-1995)</td></tr><tr><td class="cat_val">550</td><td class="cat_lbl">Other Race (1996)</td></tr><tr><td class="cat_val">560</td><td class="cat_lbl">Other Race (1997-1998)</td></tr><tr><td class="cat_val">570</td><td class="cat_lbl">Other Race (1999-2002)</td></tr><tr><td class="cat_val">580</td><td class="cat_lbl">Primary Race not releasable</td></tr><tr><td class="cat_val">600</td><td class="cat_lbl">Multiple Race, No Primary Race Selected</td></tr><tr><td class="cat_val">610</td><td class="cat_lbl">Multiple Race, including Asian, excluding Black and White</td></tr><tr><td class="cat_val">611</td><td class="cat_lbl">Multiple Race, including Asian and Black, excluding White</td></tr><tr><td class="cat_val">612</td><td class="cat_lbl">Multiple Race, including Asian and White, excluding Black</td></tr><tr><td class="cat_val">613</td><td class="cat_lbl">Multiple Race, including Black, excluding Asian and White</td></tr><tr><td class="cat_val">614</td><td class="cat_lbl">Multiple Race, including Black and White, excluding Asian</td></tr><tr><td class="cat_val">615</td><td class="cat_lbl">Multiple Race, including White, excluding Asian and Black</td></tr><tr><td class="cat_val">616</td><td class="cat_lbl">Multiple Race, including Asian, White, and Black</td></tr><tr><td class="cat_val">617</td><td class="cat_lbl">Multiple Race, excluding Asian, White, and Black</td></tr><tr><td class="cat_val">900</td><td class="cat_lbl">Unknown</td></tr><tr><td class="cat_val">970</td><td class="cat_lbl">Unknown-refused</td></tr><tr><td class="cat_val">980</td><td class="cat_lbl">Unknown-not ascertained</td></tr><tr><td class="cat_val">990</td><td class="cat_lbl">Unknown (1997 forward: Don't know)</td></tr></tbody></table>
+```
+Variable: "MARSTAT"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">MARSTAT</td></tr><tr><td class="label">Label:</td><td class="value">Legal marital status</td></tr><tr><td class="label">Variable Text:</td><td class="value">For persons age 16 and older, MARSTAT reports the person's legal marital status.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">103</td></tr><tr><td class="label">End Position:</td><td class="value">104</td></tr><tr><td class="label">Width:</td><td class="value">2</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
+<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">00</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">10</td><td class="cat_lbl">Married</td></tr><tr><td class="cat_val">20</td><td class="cat_lbl">Widowed</td></tr><tr><td class="cat_val">30</td><td class="cat_lbl">Divorced</td></tr><tr><td class="cat_val">40</td><td class="cat_lbl">Separated</td></tr><tr><td class="cat_val">50</td><td class="cat_lbl">Never married</td></tr><tr><td class="cat_val">99</td><td class="cat_lbl">Unknown marital status</td></tr></tbody></table>
+```
+Variable: "EDUC"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">EDUC</td></tr><tr><td class="label">Label:</td><td class="value">Educational attainment</td></tr><tr><td class="label">Variable Text:</td><td class="value">EDUC
+ reports the highest level of schooling an individual had completed, in 
+terms of completed grades for persons with less than a high school 
+degree, and in terms of degrees attained for high school graduates and 
+those with higher education. EDUC is available for all survey 
+participants age 5 and older at the time of their first MEPS interview. 
+It is an IPUMS MEPS constructed variable. <br><br>See Comparability section for more information on the construction of EDUC.<br><br>Education
+ Variables offered through IPUMS MEPS&amp;nbsp;In addition to EDUC, 
+there are several other educational attainment variables offered through
+ MEPS. Unlike EDUC, these variables are not available for every MEPS 
+year.<br><br>EDUCYR: Years of education completed (available 1996-2011 and 2014 forward)<br><br>HIDEG: Highest degree completed (available 1996-2011 and 2014 forward)<br><br>EDUYRDG: Years of education and highest degree completed (available 2011-2015)<br><br>EDRECODE: Years of education, recode (available 2011-2015)<br></td></tr><tr><td class="label">Concept:</td><td class="value">Education Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">108</td></tr><tr><td class="label">End Position:</td><td class="value">110</td></tr><tr><td class="label">Width:</td><td class="value">3</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
+<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">000</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">100</td><td class="cat_lbl">Grade 12 or less, no high school diploma or equivalent</td></tr><tr><td class="cat_val">101</td><td class="cat_lbl">Grade 8 or less (no further detail)</td></tr><tr><td class="cat_val">102</td><td class="cat_lbl">Never attended/kindergarten only</td></tr><tr><td class="cat_val">103</td><td class="cat_lbl">Grades 1-11 (no further detail)</td></tr><tr><td class="cat_val">104</td><td class="cat_lbl">Grade 1</td></tr><tr><td class="cat_val">105</td><td class="cat_lbl">Grade 2</td></tr><tr><td class="cat_val">106</td><td class="cat_lbl">Grade 3</td></tr><tr><td class="cat_val">107</td><td class="cat_lbl">Grade 4</td></tr><tr><td class="cat_val">108</td><td class="cat_lbl">Grade 5</td></tr><tr><td class="cat_val">109</td><td class="cat_lbl">Grade 6</td></tr><tr><td class="cat_val">110</td><td class="cat_lbl">Grade 7</td></tr><tr><td class="cat_val">111</td><td class="cat_lbl">Grade 8</td></tr><tr><td class="cat_val">112</td><td class="cat_lbl">Grade 9-12, no diploma (no further detail)</td></tr><tr><td class="cat_val">113</td><td class="cat_lbl">Grade 9</td></tr><tr><td class="cat_val">114</td><td class="cat_lbl">Grade 10</td></tr><tr><td class="cat_val">115</td><td class="cat_lbl">Grade 11</td></tr><tr><td class="cat_val">116</td><td class="cat_lbl">12th grade, no diploma</td></tr><tr><td class="cat_val">200</td><td class="cat_lbl">High school diploma or GED</td></tr><tr><td class="cat_val">201</td><td class="cat_lbl">High school graduate</td></tr><tr><td class="cat_val">202</td><td class="cat_lbl">GED or equivalent</td></tr><tr><td class="cat_val">300</td><td class="cat_lbl">Some college, no 4yr degree</td></tr><tr><td class="cat_val">301</td><td class="cat_lbl">Some college, no degree</td></tr><tr><td class="cat_val">302</td><td class="cat_lbl">AA degree: technical/vocational/occupational</td></tr><tr><td class="cat_val">303</td><td class="cat_lbl">AA degree: academic program</td></tr><tr><td class="cat_val">400</td><td class="cat_lbl">Bachelor's degree (BA,AB,BS,BBA)</td></tr><tr><td class="cat_val">500</td><td class="cat_lbl">Master's, Professional, or Doctoral Degree</td></tr><tr><td class="cat_val">501</td><td class="cat_lbl">Master's degree (MA,MS,Med,MBA)</td></tr><tr><td class="cat_val">502</td><td class="cat_lbl">Professional (MD,DDS,DVM,JD)</td></tr><tr><td class="cat_val">503</td><td class="cat_lbl">Doctoral degree (PhD, EdD)</td></tr><tr><td class="cat_val">504</td><td class="cat_lbl">Other degree</td></tr><tr><td class="cat_val">505</td><td class="cat_lbl">Professional School or Doctoral degree, topcoded (MD, DDS, DVM, JD, PhD, EdD)</td></tr><tr><td class="cat_val">996</td><td class="cat_lbl">No degree, years of education unknown</td></tr><tr><td class="cat_val">997</td><td class="cat_lbl">Unknown--refused</td></tr><tr><td class="cat_val">998</td><td class="cat_lbl">Unknown--not ascertained</td></tr><tr><td class="cat_val">999</td><td class="cat_lbl">Unknown--don't know</td></tr></tbody></table>
+```
+Variable: "INCTOT"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">INCTOT</td></tr><tr><td class="label">Label:</td><td class="value">Total personal income</td></tr><tr><td class="label">Variable Text:</td><td class="value">INCTOT
+ reports the sum of all person-level income for the current calendar 
+year, excluding income from tax refunds and capital gains.<br><br>INCTOT
+ includes annual earnings from wages, salaries, bonuses, tips, 
+commissions; business and farm games and losses; unemployment and 
+workers' compensation; interest and dividends; alimony, child support, 
+and other private cash transfers; private pensions, IRA withdrawals, 
+social security, and veterans payments; supplemental security income and
+ cash welfare payments from public assistance, Temporary Assistance for 
+Needy Families, and related programs; gains or losses from estates, 
+trusts, partnerships, S corporations, rent, and royalties; and a small 
+amount of "other" income. Person-level income excluded tax refunds and 
+capital gains. <br><br>Logical editing or weighted, sequential hot-deck 
+imputation was used to impute income amounts for missing values (both 
+for item non-response and for person in the full-year file who were not 
+in round 3).
+Reported income components were generally left unedited.
+<br><br>Related Variables<br>INCWAGE: annual wage and salary income of individuals<br><br>INCBUS: annual business income of individuals<br><br>INCUNEMP: annual unemployment compensation income of individuals<br><br>INCWKCOM: annual workers' compensation income of individuals<br><br>INCINT: annual interest income of individuals<br><br>INCDIVID: annual dividend income of individuals<br><br>INCRETIR: annual pension income of individuals<br><br>INCSS: annual social security income of individuals<br><br>INCTRST: annual trust/rent income of individuals<br><br>INCVET: annual veteran income of individuals<br><br>INCIRA: annual IRA income of individuals<br><br>INCALIM: annual alimony income of individuals<br><br>INCCHLD: annual child support income of individuals<br><br>INCCASH: annual other regular cash income income of individuals<br><br>INCSSI: annual SSI income of individuals<br><br>INCWELFR: annual public assistance income of individuals<br><br>INCOTH: annual other income of individuals<br></td></tr><tr><td class="label">Concept:</td><td class="value">Total Income and Earnings Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">111</td></tr><tr><td class="label">End Position:</td><td class="value">119</td></tr><tr><td class="label">Width:</td><td class="value">9</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">2</td></tr><tr><td class="label">Coder Instructions:</td><td class="value">CodesINCTOT
+ is a 9-digit numeric variable with 2 implied decimals. That is, values 
+of 012345678 should be interpreted as 123456.78. The command files 
+delivered with IPUMS extracts automatically divide INCTOT by 100, so no 
+further adjustment is needed.<br><br>9999999.96:	Not in Universe<br>9999999.97:	Unknown-refused<br>9999999.98:	Unknown-not ascertained<br>9999999.99:	Unknown-don't know</td></tr></tbody></table>
+```
+Variable: "K6SUM"
+```{=html}
+<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">K6SUM</td></tr><tr><td class="label">Label:</td><td class="value">K6 score for nonspecific psychological distress: last 30 days</td></tr><tr><td class="label">Variable Text:</td><td class="value">For
+ persons eligible for the self-administered questionnaire (SAQELIG), 
+K6SUM is the summed scale value measuring nonspecific psychological 
+distress over a 30-day recall period. This scale, developed by Ronald C.
+ Kessler and known as the Kessler 6 Scale (K6), asks about six 
+manifestations of nonspecific psychological distress.<br><br>Kessler 
+recommends scoring the scale by assigning 0 to 4 points for each of the 
+six questions, based on the reported frequency of the feelings (i.e., 0 
+for "none of the time"; 1 for "a little of the time"; 2 for "some of the
+ time"; 3 for "most of the time"; and 4 for "all of the time"). The 
+range for summed responses on the K6 Scale is thus 0 to 24, with 0 
+suggesting the lowest level of nonspecific psychological distress, and 
+24 suggesting the highest level of nonspecific psychological distress. 
+According to the scoring criteria proposed by Kessler, persons with a 
+score of 13 or greater are likely to be experiencing severe mental 
+illness.<br><br>Kessler's instrument asks how often, during the past 30 days, the respondent felt:<br><br>So sad that nothing could cheer you up?
+(ASAD)<br><br>Nervous?
+(ANERVOUS)<br><br>Restless or fidgety?
+(ARESTLESS)<br><br>Hopeless?
+(AHOPELESS)<br><br>That everything was an effort?
+(AEFFORT)<br><br>Worthless?
+(AWORTHLESS)<br>As noted above, acceptable responses fell into five categories, ranging from "none of the time" to "all of the time."</td></tr><tr><td class="label">Concept:</td><td class="value">Adult Mental Health Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">120</td></tr><tr><td class="label">End Position:</td><td class="value">121</td></tr><tr><td class="label">Width:</td><td class="value">2</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
+<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">96</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">98</td><td class="cat_lbl">Unknown-not ascertained</td></tr></tbody></table>
+```
 # Model details {#sec-model-details}

### 2024-10-22 paper/paper.qmd
Update paper.qmd
Updated #Introduction to include an overview of the paper's scope, variables used for training the model, and a short paragraph on why  models like this matter.

@@
 
-Overview paragraph
+Mental health issues, particularly depression, have emerged as a significant public health concern globally.
+Among other factors influencing mental health, [gender and education] stand out as critical determinants.
+Recent studies have also indicated that educational attainment may play a protective role against depression, suggesting that individuals with higher education levels experience lower rates of mental health issues.
+Conversely, gender disparities in depression prevalence remain pronounced, with research consistently showing that women report higher rates of depression than their male counterparts, even when controlling for education levels.
+Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression?
+Are women disproportionately affected by depression compared to men at the same educational level?
+Can we develop a model that predicts the likelihood of a [KSUM score of 13 or higher] given certain characteristics of an individual [please list the characteristics we used here, and what we called them in your data or graphs]?
+To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, income, and mental health outcomes [please cite the IPUMs dataset here, with a footnote on how you selected your filters].
+
+This study employs a statistical approach in R  to examine the relationships between gender, educational attainment, income and reported cases of mental depression.
+By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+
+The remainder of this paper is structured as follows: Section 2 presents the data and methodology, detailing the statistical techniques used for analysis.
+Section 3 discusses the modeling strategy and expected outcomes, followed by a presentation of results in Section 4.
+Finally, Section 5 engages in a critical discussion of our findings, addressing their implications for future research and policy.
 
@@
 Why it matters paragraph
+Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health.
+They allow researchers and policymakers spot trends and make smarter decisions.
+By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
+Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health.
+In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.
 

### 2024-10-22 paper/paper.qmd
Introduction, data section are complete.

--- 2024-10-21 paper/paper.qmd
@@
 library(here)
+library(ggplot2)
 filtered_df_2018 <- read.csv(here::here("data/02-analysis_data/model_data_2018.csv"))
 filtered_df_2021 <- read.csv(here::here("data/02-analysis_data/model_data_2021.csv"))
+
+labeled_df_reduced_2018 <- filtered_df_2018 |> mutate(
+  outcome = if_else(outcome==1, "Yes", "No"),
+  outcome = as_factor(outcome),
+  sex = if_else(sex==1,"Male", "Female"),
+  sex = as_factor(sex),
+  marital_status = case_when(
+    marstat==10~"Married",
+    marstat==20~"Widowed",
+    marstat==30~"Divorced",
+    marstat==40~"Separated",
+    marstat==50~"Never married"
+  ),
+  marital_status = factor(marital_status, levels=c("Married","Widowed","Divorced","Separated","Never married")),
+  race= case_when(
+    racea==100~"White",
+    racea==200~"Black/African-American",
+    racea==300~"Aleut, Alaskan Native, or American Indian",
+    racea==310~"Alaskan Native or American Indian",
+    racea==320~"Alaskan Native/Eskimo",
+    racea==330~"Aleut",
+    racea==340~"American Indian",
+    racea==350~"American Indian or Alaskan Native and any other group",
+    racea==400~"Asian or Pacific Islander",
+    racea==410~"Asian",
+    racea==411~"Chinese",
+    racea==412~"Filipino",
+    racea==413~"Korean",
+    racea==414~"Vietnamese",
+    racea==415~"Japanese",
+    racea==416~"Asian Indian",
+    racea==420~"Pacific Islander",
+    racea==421~"Hawaiian",
+    racea==422~"Samoan",
+    racea==423~"Guamanian",
+    racea==430~"Other Asian or Pacific Islander",
+    racea==431~"Other Asian or Pacific Islander (1992-1995)",
+    racea==432~"Other Asian or Pacific Islander (1996)",
+    racea==433~"Other Asian or Pacific Islander (1997-1998)",
+    racea==434~"Other Asian (1999 forward)",
+    racea==500~"Other Race",
+    racea==510~"Other Race (1963-1977)",
+    racea==520~"Other Race (1978)",
+    racea==530~"Other Race (1979-1991)",
+    racea==540~"Other Race (1992-1995)",
+    racea==550~"Other Race (1996)",
+    racea==560~"Other Race (1997-1998)",
+    racea==570~"Other Race (1999-2002)",
+    racea==580~"Primary Race not releasable",
+    racea==600~"Multiple Race, No Primary Race Selected",
+    racea==610~"Multiple Race, including Asian, excluding Black and White",
+    racea==611~"Multiple Race, including Asian and Black, excluding White",
+    racea==612~"Multiple Race, including Asian and White, excluding Black",
+    racea==613~"Multiple Race, including Black, excluding Asian and White",
+    racea==614~"Multiple Race, including Black and White, excluding Asian",
+    racea==615~"Multiple Race, including White, excluding Asian and Black",
+    racea==616~"Multiple Race, including Asian, White, and Black",
+    racea==617~"Multiple Race, excluding Asian, White, and Black"
+  ),
+  race = factor(race, 
+                levels=c("White", "Black/African-American", "Aleut, Alaskan Native, or American Indian", "Alaskan Native or American Indian", "Alaskan Native/Eskimo", "Aleut", "American Indian", "American Indian or Alaskan Native and any other group", "Asian or Pacific Islander", "Asian", "Chinese", "Filipino", "Korean", "Vietnamese", "Japanese", "Asian Indian", "Pacific Islander", "Hawaiian", "Samoan", "Guamanian", "Other Asian or Pacific Islander", "Other Asian or Pacific Islander (1992-1995)", "Other Asian or Pacific Islander (1996)", "Other Asian or Pacific Islander (1997-1998)", "Other Asian (1999 forward)", "Other Race", "Other Race (1963-1977)", "Other Race (1978)", "Other Race (1979-1991)", "Other Race (1992-1995)", "Other Race (1996)", "Other Race (1997-1998)", "Other Race (1999-2002)", "Primary Race not releasable", "Multiple Race, No Primary Race Selected", "Multiple Race, including Asian, excluding Black and White", "Multiple Race, including Asian and Black, excluding White", "Multiple Race, including Asian and White, excluding Black", "Multiple Race, including Black, excluding Asian and White", "Multiple Race, including Black and White, excluding Asian", "Multiple Race, including White, excluding Asian and Black", "Multiple Race, including Asian, White, and Black", "Multiple Race, excluding Asian, White, and Black"
+                )),
+  education = case_when(
+    educ==100~"Grade 12 or less",
+    educ==101~"Grade 12 or less",
+    educ==102~"Never attended/kindergarten only",
+    educ==103~"Grade 12 or less",
+    educ==104~"Grade 12 or less",
+    educ==105~"Grade 12 or less",
+    educ==106~"Grade 12 or less",
+    educ==107~"Grade 12 or less",
+    educ==108~"Grade 12 or less",
+    educ==109~"Grade 12 or less",
+    educ==110~"Grade 12 or less",
+    educ==111~"Grade 12 or less",
+    educ==112~"Grade 12 or less",
+    educ==113~"Grade 12 or less",
+    educ==114~"Grade 12 or less",
+    educ==115~"Grade 12 or less",
+    educ==116~"Grade 12 or less",
+    educ==200~"High school",
+    educ==201~"High school",
+    educ==202~"GED or equivalent",
+    educ==300~"Some college",
+    educ==301~"Some college",
+    educ==302~"AA degree",
+    educ==303~"AA degree",
+    educ==400~"Bachelor's",
+    educ==501~"Master's",
+    educ==502~"Professional",
+    educ==503~"Doctoral"
+  ),
+  education = factor(
+    education,
+    levels=c(
+      "Grade 12 or less","Never attended/kindergarten only","High school","GED or equivalent","Some college","AA degree","Bachelor's","Master's","Professional","Doctoral"
+    )
+  )
+)
+
+labeled_df_reduced_2021 <- filtered_df_2021 |> mutate(
+  outcome = if_else(outcome==1, "Yes", "No"),
+  outcome = as_factor(outcome),
+  sex = if_else(sex==1,"Male", "Female"),
+  sex = as_factor(sex),
+  marital_status = case_when(
+    marstat==10~"Married",
+    marstat==20~"Widowed",
+    marstat==30~"Divorced",
+    marstat==40~"Separated",
+    marstat==50~"Never married"
+  ),
+  marital_status = factor(marital_status, levels=c("Married","Widowed","Divorced","Separated","Never married")),
+  race= case_when(
+    racea==100~"White",
+    racea==200~"Black/African-American",
+    racea==300~"Aleut, Alaskan Native, or American Indian",
+    racea==310~"Alaskan Native or American Indian",
+    racea==320~"Alaskan Native/Eskimo",
+    racea==330~"Aleut",
+    racea==340~"American Indian",
+    racea==350~"American Indian or Alaskan Native and any other group",
+    racea==400~"Asian or Pacific Islander",
+    racea==410~"Asian",
+    racea==411~"Chinese",
+    racea==412~"Filipino",
+    racea==413~"Korean",
+    racea==414~"Vietnamese",
+    racea==415~"Japanese",
+    racea==416~"Asian Indian",
+    racea==420~"Pacific Islander",
+    racea==421~"Hawaiian",
+    racea==422~"Samoan",
+    racea==423~"Guamanian",
+    racea==430~"Other Asian or Pacific Islander",
+    racea==431~"Other Asian or Pacific Islander (1992-1995)",
+    racea==432~"Other Asian or Pacific Islander (1996)",
+    racea==433~"Other Asian or Pacific Islander (1997-1998)",
+    racea==434~"Other Asian (1999 forward)",
+    racea==500~"Other Race",
+    racea==510~"Other Race (1963-1977)",
+    racea==520~"Other Race (1978)",
+    racea==530~"Other Race (1979-1991)",
+    racea==540~"Other Race (1992-1995)",
+    racea==550~"Other Race (1996)",
+    racea==560~"Other Race (1997-1998)",
+    racea==570~"Other Race (1999-2002)",
+    racea==580~"Primary Race not releasable",
+    racea==600~"Multiple Race, No Primary Race Selected",
+    racea==610~"Multiple Race, including Asian, excluding Black and White",
+    racea==611~"Multiple Race, including Asian and Black, excluding White",
+    racea==612~"Multiple Race, including Asian and White, excluding Black",
+    racea==613~"Multiple Race, including Black, excluding Asian and White",
+    racea==614~"Multiple Race, including Black and White, excluding Asian",
+    racea==615~"Multiple Race, including White, excluding Asian and Black",
+    racea==616~"Multiple Race, including Asian, White, and Black",
+    racea==617~"Multiple Race, excluding Asian, White, and Black"
+  ),
+  race = factor(race, 
+                levels=c("White", "Black/African-American", "Aleut, Alaskan Native, or American Indian", "Alaskan Native or American Indian", "Alaskan Native/Eskimo", "Aleut", "American Indian", "American Indian or Alaskan Native and any other group", "Asian or Pacific Islander", "Asian", "Chinese", "Filipino", "Korean", "Vietnamese", "Japanese", "Asian Indian", "Pacific Islander", "Hawaiian", "Samoan", "Guamanian", "Other Asian or Pacific Islander", "Other Asian or Pacific Islander (1992-1995)", "Other Asian or Pacific Islander (1996)", "Other Asian or Pacific Islander (1997-1998)", "Other Asian (1999 forward)", "Other Race", "Other Race (1963-1977)", "Other Race (1978)", "Other Race (1979-1991)", "Other Race (1992-1995)", "Other Race (1996)", "Other Race (1997-1998)", "Other Race (1999-2002)", "Primary Race not releasable", "Multiple Race, No Primary Race Selected", "Multiple Race, including Asian, excluding Black and White", "Multiple Race, including Asian and Black, excluding White", "Multiple Race, including Asian and White, excluding Black", "Multiple Race, including Black, excluding Asian and White", "Multiple Race, including Black and White, excluding Asian", "Multiple Race, including White, excluding Asian and Black", "Multiple Race, including Asian, White, and Black", "Multiple Race, excluding Asian, White, and Black"
+                )),
+  education = case_when(
+    educ==100~"Grade 12 or less",
+    educ==101~"Grade 12 or less",
+    educ==102~"Never attended/kindergarten only",
+    educ==103~"Grade 12 or less",
+    educ==104~"Grade 12 or less",
+    educ==105~"Grade 12 or less",
+    educ==106~"Grade 12 or less",
+    educ==107~"Grade 12 or less",
+    educ==108~"Grade 12 or less",
+    educ==109~"Grade 12 or less",
+    educ==110~"Grade 12 or less",
+    educ==111~"Grade 12 or less",
+    educ==112~"Grade 12 or less",
+    educ==113~"Grade 12 or less",
+    educ==114~"Grade 12 or less",
+    educ==115~"Grade 12 or less",
+    educ==116~"Grade 12 or less",
+    educ==200~"High school",
+    educ==201~"High school",
+    educ==202~"GED or equivalent",
+    educ==300~"Some college",
+    educ==301~"Some college",
+    educ==302~"AA degree",
+    educ==303~"AA degree",
+    educ==400~"Bachelor's",
+    educ==501~"Master's",
+    educ==502~"Professional",
+    educ==503~"Doctoral"
+  ),
+  education = factor(
+    education,
+    levels=c(
+      "Grade 12 or less","Never attended/kindergarten only","High school","GED or equivalent","Some college","AA degree","Bachelor's","Master's","Professional","Doctoral"
+    )
+  )
+)
 ```
@@
 
-Overview paragraph
+Mental health issues, particularly depression, have emerged as a significant public health concern globally.
+Among other factors influencing mental health, gender and education stand out as critical determinants.
+Recent studies have also indicated that educational attainment may play a protective role against depression, suggesting that individuals with higher education levels experience lower rates of mental health issues.
+Conversely, gender disparities in depression prevalence remain pronounced, with research consistently showing that women report higher rates of depression than their male counterparts, even when controlling for education levels.
 
-Estimand paragraph
+Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression?
+Are women disproportionately affected by depression compared to men at the same educational level?
+Can we develop a model that predicts the likelihood of psychological distress given certain characteristics of an individual such as sex, education or age?
+To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, and mental health outcomes.
 
-Results paragraph
+This study employs a statistical approach in R  to examine the relationships between gender and educational attainment and reported cases of mental depression.
+By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows: @sec-data presents the data and methodology, detailing the statistical techniques used for analysis.
+@sec-model-desc discusses the modeling strategy and expected outcomes, followed by a presentation of results in @sec-result.
+Finally, @sec-discussion engages in a critical discussion of our findings, addressing their implications for future research and policy.
 
@@
 
-```{r}
-#| label: fig-education
-#| fig-cap: Education
-#| warning: false
-#| message: false
-#| echo: false
-# labeled_df |>
-#   ggplot(aes(x = label_education, fill = label_outcome)) +
-#   stat_count(position = "dodge") +
-#   facet_wrap(facets = vars(label_sex)) +
-#   theme_minimal() +
-#   labs(
-#     x = "Highest education",
-#     y = "Number of respondents",
-#     fill = "Mental Depression"
-#   ) +
-#   coord_flip() +
-#   scale_fill_brewer(palette = "Set2") +
-#   theme(legend.position = "bottom")
-```
-
 ## Overview
-We used the Medical Expenditure Panel Survey dataset from the Integrated Public Use Microdata Series, which contains microdata from different time points related to U.S.
+We used the Medical Expenditure Panel Survey(MEPS) dataset from the Integrated Public Use Microdata Series, which contains microdata from different time points related to U.S.
 healthcare expenditures [@blewett_ipums_2023].
@@
 Our dataset contains age, sex, marital status, race, highest level of education attainment, total personal income, and K6 score [@kessler_kessler_2012].All the variable definitions can be found in the Appendix [-@sec-appendix-data-details].
-
-Our goal was to identify people, based on gender and education, who are suffering from non-specific psychological distress based on the Kessler scale [@kessler_kessler_2012] before and after the COVID-19 pandemic.
 
@@
 
-Research consistently shows that women report higher rates of depression than men (CITE).
+Research consistently shows that women report higher rates of depression than men [@van_de_velde_gender_2010].
 Biological, social, and psychological factors contribute to this disparity.
-Women may experience unique stressors, such as role strain from balancing work and family obligations, as well as being subject to gender-based discrimination (CITE).
+Women may experience unique stressors, such as role strain from balancing work and family obligations, as well as being subject to gender-based discrimination.
+Townsend et al.
+see this as, "competing demands between work and the gendered expectations of domestic labor for women", which can "result in relatively more work–family conflict." Plus, they note that "this tension between work and family tends to increase among women over their lifetime" [@townsend_holding_2024].
 These pressures can exacerbate mental health issues, leading to higher rates of depression among women, particularly those  in lower socioeconomic or educational strata.
@@
 Education is often seen as a protective factor against mental health issues.
-Studies suggest that higher levels of education correlate with improved mental health outcomes (CITE).
-Educated individuals tend to have better access to mental health resources, higher socioeconomic status, and greater job security, all of which contribute to lower rates of depression (CITE).
+Studies have mixed results when studying if higher levels of education correlate with improved mental health outcomes [@tabor_mental_2021].
+Educated individuals do, however, tend to have better access to mental health resources, higher socioeconomic status, and greater job security, all of which contribute to lower rates of depression.
+Indeed, socioeconomic factor "is a consistent and reliable predictor of a vast array of outcomes across the life span, including physical and psychological health." [@noauthor_education_nodate].
 Conversely, those with lower education levels may face more financial strain, reduced employment opportunities, and limited access to healthcare, increasing their vulnerability to depression.
@@
 Despite the protective nature of education, the gender gap in mental health persists even at higher levels of educational attainment.
-Some studies suggest that women with high education levels still experience depression due to external stressors, including workplace discrimination and societal expectations (CITE).
-Additionally, the mental health benefits of education may differ between genders, with men benefiting more from the social status and economic opportunities that education provides, while women may continue to face challenges related to gender roles (cite).
+Some studies suggest that women with high education levels still experience depression due to external stressors, including workplace discrimination and societal expectations.
+Acording to Ross et.al., "Academics report stress and waning resilience, fatigue and exhaustion, and a destabilization of the work-life balance.
+Furthermore, these impacts are unequally experienced by women with children and those with caring responsibilities" [@ross_stress_2023].
+Additionally, the mental health benefits of education may differ between genders, with men benefiting more from the social status and economic opportunities that education provides, while women may continue to face challenges related to gender roles.
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The MEPS survey collects participants' K6 score for nonspecific psychological distress.
+The K6 score is a summed scale value, also known as the Kessler 6 Scale, which considers six manifestations of nonspecific psychological distress.
+A score of 13 or higher indicates likely severe mental illness [@kessler_kessler_2012].
+The variable called K6SUM represents the score in our dataset [-@sec-appendix-data-details].
+We constructed a new variable called "outcome," where 0 represents the score below 13 and 1 represents the score above 13.
+We considered zero as a person who is not experiencing psychological distress, denoted as "No," and one as a person experiencing severe psychological distress, denoted as "Yes." @fig-respondent-outcome represents the number of respondents we have in the dataset who were suffering from psychological distress.
+In 2018, 601 respondents were suffering severe psychological distress, where the number of total respondents was 15716.
+In 2021, 573 respondents were suffering from nonspecific psychological distress over the 30-day recall period, where the total number of respondents was 12685.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-respondent-outcome
+#| fig-cap: Number of Respondents Suffering Severe Psychological Distress
+#| fig-subcap: 
+#|  - "2018"
+#|  - "2021"
+#| layout-ncol: 2
 #| echo: false
+#| warning: false
+max_count <- max(table(labeled_df_reduced_2018$outcome)) + 1000
+labeled_df_reduced_2018 |> 
+  ggplot(aes(x = outcome)) +
+  geom_histogram(stat = "count", fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  geom_text(
+    stat = "count", 
+    color = "black",
+    aes(label = after_stat(count)), 
+    vjust = -0.6,  # Adjust vertical position of labels
+    size = 5       # Adjust size of labels
+  ) +
+  labs(title = "", 
+       x = "Outcome", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 20, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 16),               # X-axis label size
+    axis.title.y = element_text(size = 16),               # Y-axis label size
+    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+  )
 
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = "none") +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
+max_count <- max(table(labeled_df_reduced_2021$outcome)) + 1000
+labeled_df_reduced_2021 |> 
+  ggplot(aes(x = outcome)) +
+  geom_histogram(stat = "count", fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  geom_text(
+    stat = "count", 
+    color = "black",
+    aes(label = after_stat(count)), 
+    vjust = -0.3,  # Adjust vertical position of labels
+    size = 5       # Adjust size of labels
+  ) +
+  labs(title = "", 
+       x = "Outcome", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 20, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 16),               # X-axis label size
+    axis.title.y = element_text(size = 16),               # Y-axis label size
+    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+  )
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-# analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-# 
-# analysis_data |> 
-#   ggplot(aes(x = width, y = length)) +
-#   geom_point(alpha = 0.8) +
-#   theme_minimal() +
-#   labs(x = "Wing width (mm)",
-#        y = "Wing length (mm)")
-```
-
-Talk way more about it.
+\newpage
 
@@
 
-Add graphs, tables and text.
+We discuss the predictor variables in this section.
+We primarily use sex and education variables for our analysis.
+However, we also discuss marital status, race, and total income to understand our dataset better.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+### Sex
 
-# Model
+Our dataset has 8209 female respondents and 7507 male respondents for 2018.
+In 2021, the total number of respondents is lower compared to 2018.
+We have 6784 female respondents and 5901 male respondents for 2021 [@fig-sex].
+
+```{r}
+#| label: fig-sex
+#| fig-cap: Distribution of Respondents by Sex
+#| fig-subcap: 
+#|  - "2018"
+#|  - "2021"
+#| layout-ncol: 2
+#| echo: false
+#| warning: false
+
+max_count <- max(table(labeled_df_reduced_2018$sex)) + 1000
+labeled_df_reduced_2018 |> 
+  ggplot(aes(x = sex)) +
+  geom_histogram(stat = "count", fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  geom_text(
+    stat = "count", 
+    color = "black",
+    aes(label = after_stat(count)), 
+    vjust = -0.3,  # Adjust vertical position of labels
+    size = 5       # Adjust size of labels
+  ) +
+  labs(title = "Sex", 
+       x = "Sex", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 20, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 16),               # X-axis label size
+    axis.title.y = element_text(size = 16),               # Y-axis label size
+    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+  )
+max_count <- max(table(labeled_df_reduced_2021$sex)) + 1000
+labeled_df_reduced_2021 |> 
+  ggplot(aes(x = sex)) +
+  geom_histogram(stat = "count", fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  geom_text(
+    stat = "count", 
+    color = "black",
+    aes(label = after_stat(count)), 
+    vjust = -0.3,  # Adjust vertical position of labels
+    size = 5       # Adjust size of labels
+  ) +
+  labs(title = "Sex", 
+       x = "Sex", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 20, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 16),               # X-axis label size
+    axis.title.y = element_text(size = 16),               # Y-axis label size
+    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+  )
+```
+
+### Education
+
+There are 36 categories in the education variable.
+Five of them represent the missing value or refused to answer.
+We discarded these values.
+We also discarded the 500, 504, and 505 codes as they represent overlap (e.g., 500 represents a Master's, Professional, or Doctoral Degree).
+We also merge the categories of education.
+Details are available in the @sec-appendix-education-recoding.
+
+@fig-education-2018 and @fig-education-2021 depict the overall education status of our respondents.
+For both 2018 and 2022, a majority of respondents have at least a high school degree, followed by a bachelor's degree.
+```{r}
+#| label: fig-education-2018
+#| fig-cap: Distribution of Respondents by Education (2018)
+#| echo: false
+#| warning: false
+max_count <- max(table(labeled_df_reduced_2018$education)) + 1000
+labeled_df_reduced_2018 |> 
+  ggplot(aes(x = education)) +
+  geom_histogram(stat = "count", binwidth = 5, fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  geom_text(
+    stat = "count", 
+    color = "black",
+    aes(label = after_stat(count)), 
+    vjust = -0.3,  # Adjust vertical position of labels
+    size = 3       # Adjust size of labels
+  ) +
+  labs(title = "Education", 
+       x = "Education", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 15, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 10),               # X-axis label size
+    axis.title.y = element_text(size = 10),               # Y-axis label size
+    axis.text = element_text(size = 10),              # Axis text size (ticks)
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+```
+
+```{r}
+#| label: fig-education-2021
+#| fig-cap: Distribution of Respondents by Education (2021)
+#| echo: false
+#| warning: false
+
+max_count <- max(table(labeled_df_reduced_2021$education)) + 1000
+labeled_df_reduced_2021 |> 
+  ggplot(aes(x = education)) +
+  geom_histogram(stat = "count", binwidth = 5, fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  geom_text(
+    stat = "count", 
+    color = "black",
+    aes(label = after_stat(count)), 
+    vjust = -0.3,  # Adjust vertical position of labels
+    size = 3       # Adjust size of labels
+  ) +
+  labs(title = "Education", 
+       x = "Education", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 15, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 10),               # X-axis label size
+    axis.title.y = element_text(size = 10),               # Y-axis label size
+    axis.text = element_text(size = 10),              # Axis text size (ticks)
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+```
+\newpage
+
+### Age
+
+@fig-age The histogram demonstrates the overview of the age data distribution, depicting the high and low frequencies.
+The highest concentration of respondents is above the age of 60.
+Our dataset has a relatively low representation of the younger population.
+
+```{r}
+#| label: fig-age
+#| fig-cap: Distribution of Respondents by Education (2021)
+#| echo: false
+#| warning: false
+
+max_count <- max(table(labeled_df_reduced_2021$age)) + 1000
+labeled_df_reduced_2021 |> 
+  ggplot(aes(x = age)) +
+  geom_histogram(fill = "#013065", color = "white") +
+  scale_y_continuous(limits = c(0, max_count)) +
+  labs(title = "Age", 
+       x = "Age", 
+       y = "Frequency") +
+  theme_light()+
+  theme(
+    plot.title = element_text(size = 15, face = "bold"),  # Title font size
+    axis.title.x = element_text(size = 10),               # X-axis label size
+    axis.title.y = element_text(size = 10),               # Y-axis label size
+    axis.text = element_text(size = 10),              # Axis text size (ticks)
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+```
+\newpage
+
+# Model {#sec-model-desc}
 
@@
 
-# Results
+# Results {#sec-result}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 ```
+
+## Education Recoding {#sec-appendix-education-recoding}
+
+100~"Grade 12 or less",
+101~"Grade 12 or less",
+102~"Never attended/kindergarten only",
+103~"Grade 12 or less",
+104~"Grade 12 or less",
+105~"Grade 12 or less",
+106~"Grade 12 or less",
+107~"Grade 12 or less",
+108~"Grade 12 or less",
+109~"Grade 12 or less",
+110~"Grade 12 or less",
+111~"Grade 12 or less",
+112~"Grade 12 or less",
+113~"Grade 12 or less",
+114~"Grade 12 or less",
+115~"Grade 12 or less",
+116~"Grade 12 or less",
+200~"High school",
+201~"High school",
+202~"GED or equivalent",
+300~"Some college",
+301~"Some college",
+302~"AA degree",
+303~"AA degree",
+400~"Bachelor's",
+501~"Master's",
+502~"Professional",
+503~"Doctoral"
+
 # Model details {#sec-model-details}

### 2024-10-23 paper/paper.qmd

@@
 Mental health issues, particularly depression, have emerged as a significant public health concern globally.
-Among other factors influencing mental health, gender and education stand out as critical determinants.
+Among other factors influencing mental health, [gender and education] stand out as critical determinants.
 Recent studies have also indicated that educational attainment may play a protective role against depression, suggesting that individuals with higher education levels experience lower rates of mental health issues.
 Conversely, gender disparities in depression prevalence remain pronounced, with research consistently showing that women report higher rates of depression than their male counterparts, even when controlling for education levels.
+Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression?
+Are women disproportionately affected by depression compared to men at the same educational level?
+Can we develop a model that predicts the likelihood of a [KSUM score of 13 or higher] given certain characteristics of an individual [please list the characteristics we used here, and what we called them in your data or graphs]?
+To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, income, and mental health outcomes [please cite the IPUMs dataset here, with a footnote on how you selected your filters].
+
+This study employs a statistical approach in R to examine the relationships between gender, educational attainment, income and reported cases of mental depression.
+By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+
+The remainder of this paper is structured as follows: Section 2 presents the data and methodology, detailing the statistical techniques used for analysis.
+Section 3 discusses the modeling strategy and expected outcomes, followed by a presentation of results in Section 4.
+Finally, Section 5 engages in a critical discussion of our findings, addressing their implications for future research and policy.
 
@@
 
-This study employs a statistical approach in R  to examine the relationships between gender and educational attainment and reported cases of mental depression.
-By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health.
+They allow researchers and policymakers spot trends and make smarter decisions.
+By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
+Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health.
+In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.
 

### 2024-10-23 paper/paper.qmd
Results and discussion section added

--- 2024-10-22 paper/paper.qmd
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+We leverage the logistic regression model to identify the association between suffering from psychological distress with sex and education.
+We set our model following the @tellingstories guidelines with the rstanarm package [@rstanarm_goodrich, @rstanarm_brilleman] in R [@citer].
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Logistic regression is a commonly used regression model when the outcome variable is binary [@tellingstories].
+For our dataset, we introduce an outcome variable, which represents whether an individual is suffering from psychological distress.
+As a predictor variable for the logistic regression, we will use the sex and education variables discussed in the @sec-data.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+For logistic regression, we are using the equation below:
 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i \sim Bern(\pi_i) \\
+logit(y_i) \sim \beta_0 + \beta_1 \times sex_i + \beta_2 \times education_i \\
+\beta_0 \sim Normal(0,2.5) \\
+\beta_1 \sim Normal(0,2.5) \\
+\beta_2 \sim Normal(0,2.5)
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Here, $y_i$ represents whether a respondent suffers from psychological distress.
+$sex_i$ represents the respondent's gender, and $education_i$ represents the maximum education attained by the respondent.
+We randomly select 5000 samples from our entire dataset to simplify the model for lower execution time.
 
@@
 
-Our results are summarized in @tbl-lrmodelresults.
+Our results are summarized in @tbl-lrmodelresults-2018 and @tbl-lrmodelresults-2021.
 
@@
 #| eval: true
-#| label: tbl-lrmodelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-# mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
-# 
-# modelsummary(
-#   list(
-#     "Mental Illness Exist" = mental_illness
-#   ),
-#   statistic = "mad"
-#   )
+#| label: tbl-lrmodelresults-2018
+#| tbl-cap: "Model Results for Year 2018"
+
+mental_illness_2018 <- readRDS(file = here::here("models/mental_illness_2018.rds"))
+
+modelsummary(
+  list(
+    "Severe Mental Illness" = mental_illness_2018
+  ),
+  statistic = "mad"
+)
 ```
@@
 ```{r}
+#| warning: false
+#| message: false
 #| echo: false
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+#| label: tbl-lrmodelresults-2021
+#| tbl-cap: "Model Results for Year 2021"
 
-# modelsummary::modelsummary(
-#   list(
-#     "First model" = first_model
-#   ),
-#   statistic = "mad",
-#   fmt = 2
-# )
+mental_illness_2021 <- readRDS(file = here::here("models/mental_illness_2021.rds"))
+modelsummary(
+  list(
+    "Severe Mental Illness" = mental_illness_2021
+  ),
+  statistic = "mad"
+)
 ```
+
+\newpage
 
@@
 
-## First discussion point {#sec-first-point}
+## Gender Differences in Depression {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+@tbl-lrmodelresults-2018 and @tbl-lrmodelresults-2021 demonstrate that the odds of being female and having psychological depression are higher compared to males.
+The findings are consistent both before and after the pandemic.
 
-## Second discussion point
+## Educational Differences
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
+We have not found any notable differences between education and psychological distress.
+Research indicates a strong connection between emotional well-being and education, as emotional learning is believed to be nurtured within the school environment and through the curriculum [@norwich_critical_2022;@billington_conditions_2022].
 
@@
 
-Weaknesses and next steps should also be included.
+We randomly selected 5000 samples for our model.
+We have not used any prediction to find out how good our model is for detecting psychological depression.
+We also have not conducted any data augmentation to improve our model result.
+Our next steps are carefully curating the dataset, introducing several data validation techniques, and increasing the sample set.
 
@@
 
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-# pp_check(first_model) +
-#   theme_classic() +
-#   theme(legend.position = "bottom")
-# 
-# posterior_vs_prior(first_model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "bottom") +
-#   coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-# 
-# plot(first_model, "trace")
-# 
-# plot(first_model, "rhat")
-```
-
 \newpage

### 2024-10-23 paper/paper.qmd
merged with main

--- 2024-10-23 paper/paper.qmd
@@
 
+This study employs a statistical approach in R to examine the relationships between gender, educational attainment, income and reported cases of mental depression.
+By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+
 Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression?
 Are women disproportionately affected by depression compared to men at the same educational level?
-Can we develop a model that predicts the likelihood of psychological distress given certain characteristics of an individual such as sex, education or age?
-To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, and mental health outcomes.
+Can we develop a model that predicts the likelihood of psychological distress [@kessler_kessler_2012] given certain characteristics of an individual such as sex, education or age?
+To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, and mental health outcomes [@blewett_ipums_2023].
 
-This study employs a statistical approach in R  to examine the relationships between gender and educational attainment and reported cases of mental depression.
-By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health.
+They allow researchers and policymakers spot trends and make smarter decisions.
+By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
+Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health.
+In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.
 

### 2024-10-23 paper/paper.qmd

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Mental health poses a significant global health threat.
+This threat existed before the pandemic, and recent studies have shown that it has become even more prevalent after the pandemic.
+Our study developed a noble method to identify severe psychological depression using survey data that can be used as an early alert system to identify nonspecific psychological distress."
 format: pdf

### 2024-10-23 paper/paper.qmd
Conclusion added by Christina.

@@
 
+# Conclusion
+
+This paper presented an overview of the educational attainment, sex, and psychological distress among participants.
+By looking at the Kessler 6 Scale for nonspecific psychological distress (K6SUM), we found complex relationships between education, sex, and mental health in the United States for our population.
+Firstly, the data revealed significant trends in educational attainment, highlighting disparities that exist among different demographic groups.
+Our constructed variable, EDUC, provided a clear understanding of individuals' highest completed education level through group lumping, which is important for identifying populations at risk of lower socioeconomic status.
+This knowledge can help inform targeted educational interventions and policies aimed at improving access to higher education, such as increasing funding for scholarships and grants for underrepresented groups, expanding community college programs, implementing outreach initiatives in underserved areas, providing mentorship opportunities, and enhancing financial literacy programs to help prospective students navigate the college application and funding processes.
+
 \newpage

### 2024-10-24 paper/paper.qmd
Update paper.qmd to include corrected title, and new subtitle.

@@
 ---
-title: "Gender and Mental Depression: Analyzing Census Data Through Statistical Methods"
+title: "Sex, education, and mental health: Analyzing American census data through statistical methods, 2018 to 2021"
+subtitle: "A pre- and post-pandemic study with IPUM's MEPs data"
 author: 

### 2024-10-24 paper/paper.qmd
Update paper.qmd to be more concise in language

@@
 
-This study employs a statistical approach in R to examine the relationships between gender, educational attainment, income and reported cases of mental depression.
-By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+This study therefore employs a statistical approach in R to examine the relationships between sex, educational attainment, income and reported cases of mental depression.
+An understanding of these dynamics contribute to a deeper understanding of how education and sex intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
 
-Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression?
+The key questions underpinning this research are: does higher education correlate with lower rates of depression?
 Are women disproportionately affected by depression compared to men at the same educational level?
@@
 
-Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health.
-They allow researchers and policymakers spot trends and make smarter decisions.
-By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
+Predictive models - even simplified ones like the one we present below - are useful for understanding complex issues like mental health.
+They allow researchers and policymakers to quickly spot trends and make better decisions.
+By looking at how factors like sex and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
 Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health.
-In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.
+Being able to predict and tackle what affects mental health is key to building healthier communities.
 

### 2024-10-24 paper/paper.qmd
Corrected line 412 and 320 to remove

@@
   ) +
-  labs(title = "Sex", 
+  labs(title = " ", 
        x = "Sex", 

### 2024-10-24 paper/paper.qmd
Update paper.qmd to correct Figure 5's caption
Update paper.qmd to change 'leverage' to 'use'

@@
 
-@fig-age The histogram demonstrates the overview of the age data distribution, depicting the high and low frequencies.
+@fig-age demonstrates the overview of the age data distribution, depicting the high and low frequencies.
 The highest concentration of respondents is above the age of 60.
@@
 
-We leverage the logistic regression model to identify the association between suffering from psychological distress with sex and education.
+We use a logistic regression model to identify the association between suffering from psychological distress with sex and education.
 We set our model following the @tellingstories guidelines with the rstanarm package [@rstanarm_goodrich, @rstanarm_brilleman] in R [@citer].

### 2024-10-25 paper/paper.qmd
Following issue has been fixed: "Quick suggestions"

--- 2024-10-24 paper/paper.qmd
@@
 ---
-title: "Sex, education, and mental health: Analyzing American census data through statistical methods, 2018 to 2021"
-subtitle: "A pre- and post-pandemic study with IPUM's MEPs data"
+title: "Gender and Mental Depression: Analyzing Census Data Through Statistical Methods"
 author: 
@@
 This threat existed before the pandemic, and recent studies have shown that it has become even more prevalent after the pandemic.
-Our study developed a noble method to identify severe psychological depression using survey data that can be used as an early alert system to identify nonspecific psychological distress."
+Our study developed a method to identify severe psychological depression using survey data that can be used as an early alert system to identify nonspecific psychological distress."
 format: pdf
@@
 
-library(boot)
-library(broom.mixed)
-library(collapse)
-library(dataverse)
-library(gutenbergr)
-library(janitor)
-library(knitr)
-library(marginaleffects)
 library(modelsummary)
 library(rstanarm)
-library(tidybayes)
 library(tidyverse)
 library(here)
-library(ggplot2)
-filtered_df_2018 <- read.csv(here::here("data/02-analysis_data/model_data_2018.csv"))
-filtered_df_2021 <- read.csv(here::here("data/02-analysis_data/model_data_2021.csv"))
+library(arrow)
+library(knitr)
 
-labeled_df_reduced_2018 <- filtered_df_2018 |> mutate(
-  outcome = if_else(outcome==1, "Yes", "No"),
-  outcome = as_factor(outcome),
-  sex = if_else(sex==1,"Male", "Female"),
-  sex = as_factor(sex),
-  marital_status = case_when(
-    marstat==10~"Married",
-    marstat==20~"Widowed",
-    marstat==30~"Divorced",
-    marstat==40~"Separated",
-    marstat==50~"Never married"
-  ),
-  marital_status = factor(marital_status, levels=c("Married","Widowed","Divorced","Separated","Never married")),
-  race= case_when(
-    racea==100~"White",
-    racea==200~"Black/African-American",
-    racea==300~"Aleut, Alaskan Native, or American Indian",
-    racea==310~"Alaskan Native or American Indian",
-    racea==320~"Alaskan Native/Eskimo",
-    racea==330~"Aleut",
-    racea==340~"American Indian",
-    racea==350~"American Indian or Alaskan Native and any other group",
-    racea==400~"Asian or Pacific Islander",
-    racea==410~"Asian",
-    racea==411~"Chinese",
-    racea==412~"Filipino",
-    racea==413~"Korean",
-    racea==414~"Vietnamese",
-    racea==415~"Japanese",
-    racea==416~"Asian Indian",
-    racea==420~"Pacific Islander",
-    racea==421~"Hawaiian",
-    racea==422~"Samoan",
-    racea==423~"Guamanian",
-    racea==430~"Other Asian or Pacific Islander",
-    racea==431~"Other Asian or Pacific Islander (1992-1995)",
-    racea==432~"Other Asian or Pacific Islander (1996)",
-    racea==433~"Other Asian or Pacific Islander (1997-1998)",
-    racea==434~"Other Asian (1999 forward)",
-    racea==500~"Other Race",
-    racea==510~"Other Race (1963-1977)",
-    racea==520~"Other Race (1978)",
-    racea==530~"Other Race (1979-1991)",
-    racea==540~"Other Race (1992-1995)",
-    racea==550~"Other Race (1996)",
-    racea==560~"Other Race (1997-1998)",
-    racea==570~"Other Race (1999-2002)",
-    racea==580~"Primary Race not releasable",
-    racea==600~"Multiple Race, No Primary Race Selected",
-    racea==610~"Multiple Race, including Asian, excluding Black and White",
-    racea==611~"Multiple Race, including Asian and Black, excluding White",
-    racea==612~"Multiple Race, including Asian and White, excluding Black",
-    racea==613~"Multiple Race, including Black, excluding Asian and White",
-    racea==614~"Multiple Race, including Black and White, excluding Asian",
-    racea==615~"Multiple Race, including White, excluding Asian and Black",
-    racea==616~"Multiple Race, including Asian, White, and Black",
-    racea==617~"Multiple Race, excluding Asian, White, and Black"
-  ),
-  race = factor(race, 
-                levels=c("White", "Black/African-American", "Aleut, Alaskan Native, or American Indian", "Alaskan Native or American Indian", "Alaskan Native/Eskimo", "Aleut", "American Indian", "American Indian or Alaskan Native and any other group", "Asian or Pacific Islander", "Asian", "Chinese", "Filipino", "Korean", "Vietnamese", "Japanese", "Asian Indian", "Pacific Islander", "Hawaiian", "Samoan", "Guamanian", "Other Asian or Pacific Islander", "Other Asian or Pacific Islander (1992-1995)", "Other Asian or Pacific Islander (1996)", "Other Asian or Pacific Islander (1997-1998)", "Other Asian (1999 forward)", "Other Race", "Other Race (1963-1977)", "Other Race (1978)", "Other Race (1979-1991)", "Other Race (1992-1995)", "Other Race (1996)", "Other Race (1997-1998)", "Other Race (1999-2002)", "Primary Race not releasable", "Multiple Race, No Primary Race Selected", "Multiple Race, including Asian, excluding Black and White", "Multiple Race, including Asian and Black, excluding White", "Multiple Race, including Asian and White, excluding Black", "Multiple Race, including Black, excluding Asian and White", "Multiple Race, including Black and White, excluding Asian", "Multiple Race, including White, excluding Asian and Black", "Multiple Race, including Asian, White, and Black", "Multiple Race, excluding Asian, White, and Black"
-                )),
-  education = case_when(
-    educ==100~"Grade 12 or less",
-    educ==101~"Grade 12 or less",
-    educ==102~"Never attended/kindergarten only",
-    educ==103~"Grade 12 or less",
-    educ==104~"Grade 12 or less",
-    educ==105~"Grade 12 or less",
-    educ==106~"Grade 12 or less",
-    educ==107~"Grade 12 or less",
-    educ==108~"Grade 12 or less",
-    educ==109~"Grade 12 or less",
-    educ==110~"Grade 12 or less",
-    educ==111~"Grade 12 or less",
-    educ==112~"Grade 12 or less",
-    educ==113~"Grade 12 or less",
-    educ==114~"Grade 12 or less",
-    educ==115~"Grade 12 or less",
-    educ==116~"Grade 12 or less",
-    educ==200~"High school",
-    educ==201~"High school",
-    educ==202~"GED or equivalent",
-    educ==300~"Some college",
-    educ==301~"Some college",
-    educ==302~"AA degree",
-    educ==303~"AA degree",
-    educ==400~"Bachelor's",
-    educ==501~"Master's",
-    educ==502~"Professional",
-    educ==503~"Doctoral"
-  ),
-  education = factor(
-    education,
-    levels=c(
-      "Grade 12 or less","Never attended/kindergarten only","High school","GED or equivalent","Some college","AA degree","Bachelor's","Master's","Professional","Doctoral"
-    )
-  )
-)
+filtered_df_2018 <- read_parquet(here::here("data/02-analysis_data/model_data_2018.parquet"))
+filtered_df_2021 <- read_parquet(here::here("data/02-analysis_data/model_data_2021.parquet"))
 
-labeled_df_reduced_2021 <- filtered_df_2021 |> mutate(
-  outcome = if_else(outcome==1, "Yes", "No"),
-  outcome = as_factor(outcome),
-  sex = if_else(sex==1,"Male", "Female"),
-  sex = as_factor(sex),
-  marital_status = case_when(
-    marstat==10~"Married",
-    marstat==20~"Widowed",
-    marstat==30~"Divorced",
-    marstat==40~"Separated",
-    marstat==50~"Never married"
-  ),
-  marital_status = factor(marital_status, levels=c("Married","Widowed","Divorced","Separated","Never married")),
-  race= case_when(
-    racea==100~"White",
-    racea==200~"Black/African-American",
-    racea==300~"Aleut, Alaskan Native, or American Indian",
-    racea==310~"Alaskan Native or American Indian",
-    racea==320~"Alaskan Native/Eskimo",
-    racea==330~"Aleut",
-    racea==340~"American Indian",
-    racea==350~"American Indian or Alaskan Native and any other group",
-    racea==400~"Asian or Pacific Islander",
-    racea==410~"Asian",
-    racea==411~"Chinese",
-    racea==412~"Filipino",
-    racea==413~"Korean",
-    racea==414~"Vietnamese",
-    racea==415~"Japanese",
-    racea==416~"Asian Indian",
-    racea==420~"Pacific Islander",
-    racea==421~"Hawaiian",
-    racea==422~"Samoan",
-    racea==423~"Guamanian",
-    racea==430~"Other Asian or Pacific Islander",
-    racea==431~"Other Asian or Pacific Islander (1992-1995)",
-    racea==432~"Other Asian or Pacific Islander (1996)",
-    racea==433~"Other Asian or Pacific Islander (1997-1998)",
-    racea==434~"Other Asian (1999 forward)",
-    racea==500~"Other Race",
-    racea==510~"Other Race (1963-1977)",
-    racea==520~"Other Race (1978)",
-    racea==530~"Other Race (1979-1991)",
-    racea==540~"Other Race (1992-1995)",
-    racea==550~"Other Race (1996)",
-    racea==560~"Other Race (1997-1998)",
-    racea==570~"Other Race (1999-2002)",
-    racea==580~"Primary Race not releasable",
-    racea==600~"Multiple Race, No Primary Race Selected",
-    racea==610~"Multiple Race, including Asian, excluding Black and White",
-    racea==611~"Multiple Race, including Asian and Black, excluding White",
-    racea==612~"Multiple Race, including Asian and White, excluding Black",
-    racea==613~"Multiple Race, including Black, excluding Asian and White",
-    racea==614~"Multiple Race, including Black and White, excluding Asian",
-    racea==615~"Multiple Race, including White, excluding Asian and Black",
-    racea==616~"Multiple Race, including Asian, White, and Black",
-    racea==617~"Multiple Race, excluding Asian, White, and Black"
-  ),
-  race = factor(race, 
-                levels=c("White", "Black/African-American", "Aleut, Alaskan Native, or American Indian", "Alaskan Native or American Indian", "Alaskan Native/Eskimo", "Aleut", "American Indian", "American Indian or Alaskan Native and any other group", "Asian or Pacific Islander", "Asian", "Chinese", "Filipino", "Korean", "Vietnamese", "Japanese", "Asian Indian", "Pacific Islander", "Hawaiian", "Samoan", "Guamanian", "Other Asian or Pacific Islander", "Other Asian or Pacific Islander (1992-1995)", "Other Asian or Pacific Islander (1996)", "Other Asian or Pacific Islander (1997-1998)", "Other Asian (1999 forward)", "Other Race", "Other Race (1963-1977)", "Other Race (1978)", "Other Race (1979-1991)", "Other Race (1992-1995)", "Other Race (1996)", "Other Race (1997-1998)", "Other Race (1999-2002)", "Primary Race not releasable", "Multiple Race, No Primary Race Selected", "Multiple Race, including Asian, excluding Black and White", "Multiple Race, including Asian and Black, excluding White", "Multiple Race, including Asian and White, excluding Black", "Multiple Race, including Black, excluding Asian and White", "Multiple Race, including Black and White, excluding Asian", "Multiple Race, including White, excluding Asian and Black", "Multiple Race, including Asian, White, and Black", "Multiple Race, excluding Asian, White, and Black"
-                )),
-  education = case_when(
-    educ==100~"Grade 12 or less",
-    educ==101~"Grade 12 or less",
-    educ==102~"Never attended/kindergarten only",
-    educ==103~"Grade 12 or less",
-    educ==104~"Grade 12 or less",
-    educ==105~"Grade 12 or less",
-    educ==106~"Grade 12 or less",
-    educ==107~"Grade 12 or less",
-    educ==108~"Grade 12 or less",
-    educ==109~"Grade 12 or less",
-    educ==110~"Grade 12 or less",
-    educ==111~"Grade 12 or less",
-    educ==112~"Grade 12 or less",
-    educ==113~"Grade 12 or less",
-    educ==114~"Grade 12 or less",
-    educ==115~"Grade 12 or less",
-    educ==116~"Grade 12 or less",
-    educ==200~"High school",
-    educ==201~"High school",
-    educ==202~"GED or equivalent",
-    educ==300~"Some college",
-    educ==301~"Some college",
-    educ==302~"AA degree",
-    educ==303~"AA degree",
-    educ==400~"Bachelor's",
-    educ==501~"Master's",
-    educ==502~"Professional",
-    educ==503~"Doctoral"
-  ),
-  education = factor(
-    education,
-    levels=c(
-      "Grade 12 or less","Never attended/kindergarten only","High school","GED or equivalent","Some college","AA degree","Bachelor's","Master's","Professional","Doctoral"
-    )
-  )
-)
+labeled_df_reduced_2018 <- filtered_df_2018
+
+labeled_df_reduced_2021 <- filtered_df_2021
 ```
@@
 
-This study therefore employs a statistical approach in R to examine the relationships between sex, educational attainment, income and reported cases of mental depression.
-An understanding of these dynamics contribute to a deeper understanding of how education and sex intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
+This study employs a statistical approach in R to examine the relationships between gender, educational attainment, income and reported cases of mental depression.
+By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.
 
-The key questions underpinning this research are: does higher education correlate with lower rates of depression?
+Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression?
 Are women disproportionately affected by depression compared to men at the same educational level?
@@
 
-Predictive models - even simplified ones like the one we present below - are useful for understanding complex issues like mental health.
-They allow researchers and policymakers to quickly spot trends and make better decisions.
-By looking at how factors like sex and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
+Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health.
+They allow researchers and policymakers spot trends and make smarter decisions.
+By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most.
 Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health.
-Being able to predict and tackle what affects mental health is key to building healthier communities.
+In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.
 
@@
 To support our analyses, we split into two datasets for 2018 and 2021.
-We also excluded the respondents whose ages are below 19 and over 80.
-After enforcing our inclusion criteria, we had 15716 respondents for 2018 and 12685 for 2021.
+We excluded the respondents whose ages are below 19 and over 80.
+After enforcing our inclusion criteria, we had 14146 respondents for 2018 and 11383 for 2021.
 To give a preview of our data, @tbl-2018 and @tbl-2021 show the first five rows for each row for 2018 and 2021, respectively.
@@
 A score of 13 or higher indicates likely severe mental illness [@kessler_kessler_2012].
-The variable called K6SUM represents the score in our dataset [-@sec-appendix-data-details].
+The variable called k6sum represents the score in our dataset [-@sec-appendix-data-details].
 We constructed a new variable called "outcome," where 0 represents the score below 13 and 1 represents the score above 13.
 We considered zero as a person who is not experiencing psychological distress, denoted as "No," and one as a person experiencing severe psychological distress, denoted as "Yes." @fig-respondent-outcome represents the number of respondents we have in the dataset who were suffering from psychological distress.
-In 2018, 601 respondents were suffering severe psychological distress, where the number of total respondents was 15716.
-In 2021, 573 respondents were suffering from nonspecific psychological distress over the 30-day recall period, where the total number of respondents was 12685.
+In 2018, 544 respondents were suffering severe psychological distress, where the number of total respondents was 14146.
+In 2021, 513 respondents were suffering from nonspecific psychological distress over the 30-day recall period, where the total number of respondents was 11383.
 
@@
     aes(label = after_stat(count)), 
-    vjust = -0.6,  # Adjust vertical position of labels
-    size = 5       # Adjust size of labels
+    vjust = -0.6,   
+    size = 5        
   ) +
@@
   theme(
-    plot.title = element_text(size = 20, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 16),               # X-axis label size
-    axis.title.y = element_text(size = 16),               # Y-axis label size
-    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+    plot.title = element_text(size = 20, face = "bold"),   
+    axis.title.x = element_text(size = 16),                
+    axis.title.y = element_text(size = 16),                
+    axis.text = element_text(size = 14)                    
   )
@@
     aes(label = after_stat(count)), 
-    vjust = -0.3,  # Adjust vertical position of labels
-    size = 5       # Adjust size of labels
+    vjust = -0.3, 
+    size = 5      
   ) +
@@
   theme(
-    plot.title = element_text(size = 20, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 16),               # X-axis label size
-    axis.title.y = element_text(size = 16),               # Y-axis label size
-    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+    plot.title = element_text(size = 20, face = "bold"),  
+    axis.title.x = element_text(size = 16),               
+    axis.title.y = element_text(size = 16),               
+    axis.text = element_text(size = 14)                   
   )
@@
 
-Our dataset has 8209 female respondents and 7507 male respondents for 2018.
+Our dataset has 7312 female respondents and 6834 male respondents for 2018.
 In 2021, the total number of respondents is lower compared to 2018.
-We have 6784 female respondents and 5901 male respondents for 2021 [@fig-sex].
+We have 6015 female respondents and 5368 male respondents for 2021 [@fig-sex].
 
@@
 #| label: fig-sex
-#| fig-cap: Distribution of Respondents by Sex
-#| fig-subcap: 
-#|  - "2018"
-#|  - "2021"
-#| layout-ncol: 2
+#| fig-cap: Distribution of Respondents by Sex (2018 and 2021)
+#| layout-ncol: 1
 #| echo: false
 #| warning: false
+# Join dataframe by sex
+count_2018 <- labeled_df_reduced_2018 %>%
+  group_by(sex) %>%
+  summarise(Count = n()) %>%
+  mutate(Year = "2018")
 
-max_count <- max(table(labeled_df_reduced_2018$sex)) + 1000
-labeled_df_reduced_2018 |> 
-  ggplot(aes(x = sex)) +
-  geom_histogram(stat = "count", fill = "#013065", color = "white") +
-  scale_y_continuous(limits = c(0, max_count)) +
-  geom_text(
-    stat = "count", 
-    color = "black",
-    aes(label = after_stat(count)), 
-    vjust = -0.3,  # Adjust vertical position of labels
-    size = 5       # Adjust size of labels
+# Count gender distribution for 2021
+count_2021 <- labeled_df_reduced_2021 %>%
+  group_by(sex) %>%
+  summarise(Count = n()) %>%
+  mutate(Year = "2021")
+
+combined <- bind_rows(count_2018, count_2021)
+max_count <-  max(combined$Count) + 1000
+ggplot(combined, aes(x = sex, y = Count, fill = Year)) +
+  geom_bar(stat = "identity", position = "dodge", width = 0.6) +
+  scale_y_continuous(limits=c(0,max_count)) + 
+    geom_text(aes(label = Count), 
+            position = position_dodge(width = 0.6),
+            vjust = -0.3,
+            size = 5) +
+  labs(
+    title = "Gender Distribution: 2018 vs 2021",
+    x = "Gender",
+    y = "Count"
   ) +
-  labs(title = " ", 
-       x = "Sex", 
-       y = "Frequency") +
-  theme_light()+
+  theme_light() +
   theme(
-    plot.title = element_text(size = 20, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 16),               # X-axis label size
-    axis.title.y = element_text(size = 16),               # Y-axis label size
-    axis.text = element_text(size = 14)                   # Axis text size (ticks)
-  )
-max_count <- max(table(labeled_df_reduced_2021$sex)) + 1000
-labeled_df_reduced_2021 |> 
-  ggplot(aes(x = sex)) +
-  geom_histogram(stat = "count", fill = "#013065", color = "white") +
-  scale_y_continuous(limits = c(0, max_count)) +
-  geom_text(
-    stat = "count", 
-    color = "black",
-    aes(label = after_stat(count)), 
-    vjust = -0.3,  # Adjust vertical position of labels
-    size = 5       # Adjust size of labels
-  ) +
-  labs(title = " ", 
-       x = "Sex", 
-       y = "Frequency") +
-  theme_light()+
-  theme(
-    plot.title = element_text(size = 20, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 16),               # X-axis label size
-    axis.title.y = element_text(size = 16),               # Y-axis label size
-    axis.text = element_text(size = 14)                   # Axis text size (ticks)
+    plot.title = element_text(size = 18, face = "bold"),
+    axis.text = element_text(size = 14),
+    axis.title = element_text(size = 16)
   )
@@
 @fig-education-2018 and @fig-education-2021 depict the overall education status of our respondents.
-For both 2018 and 2022, a majority of respondents have at least a high school degree, followed by a bachelor's degree.
+For both 2018 and 2022, a majority of respondents have at least a bachelor's, master's or PhD degree, followed by high school degree.
 ```{r}
@@
 #| warning: false
-max_count <- max(table(labeled_df_reduced_2018$education)) + 1000
-labeled_df_reduced_2018 |> 
+
+education_df_reduced_categories <- labeled_df_reduced_2018 |> 
+  select_all() |>
+  mutate(
+    education=case_when(
+      education=='Bachelor\'s'~'Bachelor\'s, Master\'s or PhD',
+      education=='Master\'s'~'Bachelor\'s, Master\'s or PhD',
+      education=='Doctoral'~'Bachelor\'s, Master\'s or PhD',
+      TRUE~education
+    )
+  )
+max_count <- max(table(education_df_reduced_categories$education)) + 1000
+education_df_reduced_categories |> 
   ggplot(aes(x = education)) +
@@
   ) +
-  labs(title = "Education", 
+  labs(title = "", 
        x = "Education", 
@@
   theme(
-    plot.title = element_text(size = 15, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 10),               # X-axis label size
-    axis.title.y = element_text(size = 10),               # Y-axis label size
-    axis.text = element_text(size = 10),              # Axis text size (ticks)
+    plot.title = element_text(size = 15, face = "bold"),  
+    axis.title.x = element_text(size = 10),               
+    axis.title.y = element_text(size = 10),               
+    axis.text = element_text(size = 10),             
     axis.text.x = element_text(angle = 45, hjust = 1)
@@
 
-max_count <- max(table(labeled_df_reduced_2021$education)) + 1000
-labeled_df_reduced_2021 |> 
+education_df_reduced_categories <- labeled_df_reduced_2021 |> 
+  select_all() |>
+  mutate(
+    education=case_when(
+      education=='Bachelor\'s'~'Bachelor\'s, Master\'s or PhD',
+      education=='Master\'s'~'Bachelor\'s, Master\'s or PhD',
+      education=='Doctoral'~'Bachelor\'s, Master\'s or PhD',
+      TRUE~education
+    )
+  )
+
+max_count <- max(table(education_df_reduced_categories$education)) + 1000
+education_df_reduced_categories |> 
   ggplot(aes(x = education)) +
@@
     aes(label = after_stat(count)), 
-    vjust = -0.3,  # Adjust vertical position of labels
-    size = 3       # Adjust size of labels
+    vjust = -0.3,  
+    size = 3       
   ) +
-  labs(title = "Education", 
+  labs(title = "", 
        x = "Education", 
@@
   theme(
-    plot.title = element_text(size = 15, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 10),               # X-axis label size
-    axis.title.y = element_text(size = 10),               # Y-axis label size
-    axis.text = element_text(size = 10),              # Axis text size (ticks)
+    plot.title = element_text(size = 15, face = "bold"),
+    axis.title.x = element_text(size = 10),        
+    axis.title.y = element_text(size = 10),       
+    axis.text = element_text(size = 10),
     axis.text.x = element_text(angle = 45, hjust = 1)
@@
 #| label: fig-age
-#| fig-cap: Distribution of Respondents by Education (2021)
+#| fig-cap: Distribution of Respondents by Age (2018, 2021)
 #| echo: false
@@
 
-max_count <- max(table(labeled_df_reduced_2021$age)) + 1000
-labeled_df_reduced_2021 |> 
+all_age <-bind_rows(labeled_df_reduced_2021,labeled_df_reduced_2018)
+max_count <- max(table(all_age$age)) + 1000
+all_age |> 
   ggplot(aes(x = age)) +
@@
   scale_y_continuous(limits = c(0, max_count)) +
-  labs(title = "Age", 
+  labs(title = "", 
        x = "Age", 
@@
   theme(
-    plot.title = element_text(size = 15, face = "bold"),  # Title font size
-    axis.title.x = element_text(size = 10),               # X-axis label size
-    axis.title.y = element_text(size = 10),               # Y-axis label size
-    axis.text = element_text(size = 10),              # Axis text size (ticks)
+    plot.title = element_text(size = 15, face = "bold"),
+    axis.title.x = element_text(size = 10),
+    axis.title.y = element_text(size = 10),
+    axis.text = element_text(size = 10),
     axis.text.x = element_text(angle = 45, hjust = 1)
@@
 
-We leverage the logistic regression model to identify the association between suffering from psychological distress with sex and education.
-We set our model following the @tellingstories guidelines with the rstanarm package [@rstanarm_goodrich, @rstanarm_brilleman] in R [@citer].
+We use the logistic regression model to identify the association between suffering from psychological distress with sex and education.
+We set our model following the @tellingstories guidelines with the rstanarm package [@rstanarm_goodrich;@rstanarm_brilleman] in R [@citeR].
 
@@
 
-We have not found any notable differences between education and psychological distress.
+We have found that psychological distress and education have a relation in terms of people who have higher education tend to feel lower psychological distress.
 Research indicates a strong connection between emotional well-being and education, as emotional learning is believed to be nurtured within the school environment and through the curriculum [@norwich_critical_2022;@billington_conditions_2022].
@@
 We have not used any prediction to find out how good our model is for detecting psychological depression.
-We also have not conducted any data augmentation to improve our model result.
 Our next steps are carefully curating the dataset, introducing several data validation techniques, and increasing the sample set.

### 2024-10-25 scripts/01-test_simulated_data.R

@@ -1,15 +1,13 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated IPUMS 
+  #MEPS dataset.
+# Author: Christina Nguyen - Marcus Barnes - Nibras Ar Rakib
+# Date: 24 October 2024
+# Contact: nibras.rakib@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
 
 #### Workspace setup ####
 library(tidyverse)
@@ -25,65 +23,34 @@ if (exists("analysis_data")) {
 
 
 #### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset 2018,2021 year
+valid_years <- c(2018, 2021)
+if (all(analysis_data$year %in% valid_years)) {
+  message("Test Passed: The 'Year' column contains only 2018 and 2021.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The 'Year' column contains invalid year.")
 }
 
 # Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+if (ncol(analysis_data) == 8) {
+  message("Test Passed: The dataset has 8 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 8 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
+# Check if the sex column has 2 unique values
+valid_sex <- c("Male","Female")
+dataset_sex <- unique(analysis_data$sex)
+if(all(valid_sex %in% dataset_sex)){
+  message("Test Passed: variable sex is valid.")
+} else{
[…]

### 2024-10-25 paper/paper.qmd
Abstract fixed

@@
 date-format: long
-abstract: "Mental health poses a significant global health threat.
-This threat existed before the pandemic, and recent studies have shown that it has become even more prevalent after the pandemic.
+abstract: "Mental health posed a significant global health threat before the pandemic, and recent studies have shown that it has become even more prevalent.
 Our study developed a method to identify severe psychological depression using survey data that can be used as an early alert system to identify nonspecific psychological distress."

### 2024-10-25 paper/paper.qmd

@@
   ) +
-  theme_light() +
   theme(

### 2024-10-25 paper/paper.qmd

@@
 This paper presented an overview of the educational attainment, sex, and psychological distress among participants.
-By looking at the Kessler 6 Scale for nonspecific psychological distress (K6SUM), we found complex relationships between education, sex, and mental health in the United States for our population.
-Firstly, the data revealed significant trends in educational attainment, highlighting disparities that exist among different demographic groups.
-Our constructed variable, EDUC, provided a clear understanding of individuals' highest completed education level through group lumping, which is important for identifying populations at risk of lower socioeconomic status.
+By looking at the Kessler 6 Scale for nonspecific psychological distress (K6SUM), we found complex relationships between education, sex, and mental health in the United States.
+Firstly, the data revealed significant trends in educational attainment, highlighting disparities that exist among different male and female.
+Our constructed variable, outcome, provided a clear understanding of individuals' highest completed education level through group lumping, which is important for identifying populations at risk of lower socioeconomic status.
 This knowledge can help inform targeted educational interventions and policies aimed at improving access to higher education, such as increasing funding for scholarships and grants for underrepresented groups, expanding community college programs, implementing outreach initiatives in underserved areas, providing mentorship opportunities, and enhancing financial literacy programs to help prospective students navigate the college application and funding processes.

### 2024-10-26 paper/paper.qmd
Adds Erickson and Kessler citations to the introduction and fixes a couple of minor typos.

@@
 Among other factors influencing mental health, gender and education stand out as critical determinants.
-Recent studies have also indicated that educational attainment may play a protective role against depression, suggesting that individuals with higher education levels experience lower rates of mental health issues.
-Conversely, gender disparities in depression prevalence remain pronounced, with research consistently showing that women report higher rates of depression than their male counterparts, even when controlling for education levels.
+Studies have also indicated that educational attainment may play a protective role against depression, suggesting that individuals with higher education levels experience lower rates of mental health issues [@erickson2016].
+Conversely, gender disparities in depression prevalence remain pronounced, with research consistently showing that women report higher rates of depression than their male counterparts, even when controlling for education levels [@kessler2006epidemiology].
 
@@
 ## Overview
-We used the Medical Expenditure Panel Survey(MEPS) dataset from the Integrated Public Use Microdata Series, which contains microdata from different time points related to U.S.
+We used the Medical Expenditure Panel Survey (MEPS) dataset from the Integrated Public Use Microdata Series, which contains microdata from different time points related to U.S.
 healthcare expenditures [@blewett_ipums_2023].
 We performed all analysis using a popular statistical language called R [@citeR].
-Our dataset contains age, sex, marital status, race, highest level of education attainment, total personal income, and K6 score [@kessler_kessler_2012].All the variable definitions can be found in the Appendix [-@sec-appendix-data-details].
+Our dataset contains age, sex, marital status, race, highest level of education attainment, total personal income, and K6 score [@kessler_kessler_2012].
+All the variable definitions can be found in the Appendix [-@sec-appendix-data-details].
 
